### 2011-05-26 pkg/paper/Paper/draft.Rnw
updated paper and draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@103 b679b2f7-6539-45d0-a605-5ddcb976cde9

\documentclass[nojss]{jss}
\usepackage{amsfonts,amstext,amsmath}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}

%% need no \usepackage{Sweave}
\setkeys{Gin}{width=\textwidth}
\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide, keep.source=TRUE}

%% for internal use
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}

\title{Structural Breaks in Inflation Dynamics within the European Monetary Union}

\author{Thomas Windberger\\Universit\"at Innsbruck \And
        Achim Zeileis\\Universit\"at Innsbruck}
\Plainauthor{Thomas Windberger, Achim Zeileis}

\Abstract{
  In this paper, we test for structural breaks in the inflation dynamics of 21 European countries. To capture
  changes in the first three moments of the inflation rates, we assumed them to be distributed according to 
  a generalized logistic distribution, which we implemented in a new \proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} 
  \citep{R} package called \pkg{glogis}~\Sexpr{packageDescription("glogis")["Version"]}. We find breaks in all but 4 countries.
  Overall, there is a convergence towards a lower inflation rate, but with an increase in variance and a change towards negative skewness.
}

\Keywords{inflation rate, structural break, EMU, generalized logistic distribution}

\Address{
  Thomas Windberger, Achim Zeileis\\
  Department of Statistics\\
  Universit\"at Innsbruck\\
  Universit\"atsstra{\ss}e 15\\
  A-6020 Innsbruck, Austria\\
  E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
}

\begin{document}

<<preliminaries>>=
## locale settings
if(Sys.getlocale() != "C") Sys.setlocale("LC_TIME", "English")

## packages
library("glogis")
library("fxregime")

## additional moment method
moments <- function(object, ...) UseMethod("moments")
moments.glogisfit <- function(object, ...) object$moments
moments.breakpoints.glogisfit <- function(object, breaks = NULL, ...)
  t(sapply(refit(object, breaks = breaks), "[[", "moments"))
@

\section{Introduction}

Ever since the beginning of the European Union, the topic of a common currency
was a controversial issue. Although the Economic and Monetary Union (EMU) is now
a fact, the discussion about the economic effects of the euro is far from being
settled. The controversial topics range from the question of whether or not the
eurozone is indeed an optimal currency area (as developed in: \citet{Mundell:1961})\footnote{
For an overview, the interested reader is referred to \citet{Dellas:2009}}, 
all the way to the very survival of the euro in light of the budgetary problems of some of its member
states. The effects of monetary unions on a number of macroeconomic indicators,
with inflation being the most important one in this context, is in the center of
an ongoing debate: this regards the question about short-run and steady-state
inflation uncertainty (in context to inflation expectations) dealt with in
\citet{cap}, or the structural convergence of the inflation rates in EU
countries, which is the topic of an investigation by \citet{palomba}.

\readme{perhaps skip this paragraph?} The way towards the EMU basically consisted of three stages: stage~I
(1990--1994) as a phase of liberalization, stage~II (1995--1998) a phase of
convergence and stage~III (1999--2001) the transition period, which ended with
the introduction of the euro as legal tender. A possible fourth stage is the
continuing integration of new member states, mostly former communist countries.
Any new state that wants to join the EMU is first obliged to fulfill the
Maastricht criteria, which, besides setting rules for government debts and
interest rates, also requires the participation in the ERM~II (Exchange Rate
Mechanism~II) for two years, while the exchange rates towards the euro are not
allowed to cross the nominal band.

\clearpage
\cite{emerson} put emphasize on the fact that a high inflation rate is also more
variable and uncertain and thus causes more relative price variability, leading
to a less efficient price mechanism.  Failing to stabilize inflation leads to
severe economic problems. Throughout the literature, there is still a
considerable degree of uncertainty as to what extent the introduction of the
euro, or monetary unions in general, affect the inflation rate. The EMU requires
a country to possibly alter a number of economic policies and institutions,
which may require significant short term adjustments.

Given that a country experienced quite volatile inflation rates, its efforts to
meet the convergence criteria are likely to lead to an alteration at least in
the mean (since this was one of the requirements) and possibly in the volatility
of their respective inflation rates as well.

However, a deeper inquiry into this problem is in need of a method able to
capture changes in inflation dynamics. We contribute to this  by
developing a new method for testing for structural breaks in the dynamics of the
inflation rates and thereby upon the effect of the EMU on a number of European
countries. The framework developed in this paper makes it possible to test for any 
significant changes of a countries inflation dynamics and thus helps interpreting 
the effects of the EMU. We focus on changes in the first three moments of these 
inflation rates, which we assume to be distributed according to a generalized logistic 
distribution (GL). The GL distribution is a suitable one due to its flexibility to 
capture nonsymmetry and because of its fatter tails. For this purpose, we develop a 
new estimation approach building on previous research by \citet{Zeileis+Hornik:2007}.

This \fixme{Jesus: sell method better, why this one and not another} framework
allows us to assess the evolution of mean, variance and skewness over time. The
skewness parameter turns out to be quite important to measure alterations in the
long run inflation dynamics, as it changes considerably over the time of the
analysis. This trend would be all but invisible using a normal distribution.  
Our estimation technique enables the simultaneous estimation of breaks in the
mean (which is important in the context of the Maastricht criteria), the
variance (which should be low for the price mechanism to be efficient) and the
skewness (where a change in skewness is evidence of a shift of the distribution
towards a new ``center of gravity'') of the inflation rates of interest. This
than can be used to search for answers regarding the question about the effect
of the EMU on the inflation dynamics of its member states or more general to
investigate the effects of external shocks.

The \fixme{Jesus: maybe comment on results here} remainder of this paper is
structured as follows: Section~\ref{sec:data} presents the data,
Section~\ref{sec:model} presents the model and the estimation techniques used,
Section~\ref{sec:slovenia} illustrates our approach using Slovenia as an
example, Section~\ref{sec:results} presents the results and
Section~\ref{sec:conclusion} concludes.

\clearpage
\section{Data}\label{sec:data}

<<data>>=
data("hicps", package = "glogis")
eu <- na.omit(hicps[, "EU"])
hicps <- hicps[, colnames(hicps) != "EU"]
@

All empirical analyses are based on seasonally adjusted inflation rates (in percent)
for \Sexpr{ncol(hicps)}~countries
(\Sexpr{gsub("UK", "United Kingdom", gsub("CzechRep", "Czech Rep", paste(colnames(hicps), collapse = ", "), fixed = TRUE), fixed = TRUE)})
in the time period from the early or mid-1990s to the end of 2010.
The original data source are seasonally unadjusted HICP series provided by the
\citet{oecd} from January~1990 (if available) to December~2010, that were transformed
to inflation rates using log-returns and subsequent seasonal adjustment
via X-12-ARIMA \citep{Findley+Monsell+Bell:1998}.

The countries in this sample can be divided into three different groups: (1)~euro
countries -- Austria, Belgium, Estonia, France, Finland, Germany, Greece,
Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia,
Spain; (2)~EU members not participating in the ERM~II -- Czech Republic,
Hungary, Poland, United Kingdom, Sweden; (3)~Denmark which stands on its own as a
member of the EU and the ERM~II, but not yet of the EMU. 
Latvia, Lithuania, Bulgaria, Romania are excluded due to data scarcity.
Cyprus and Malta are not included since they are very small economies.

Seasonal adjustment is necessary due to the harmonization of the treatment
of sales prices\footnote{Inclusion of sales prices in the HICP was demanded by
a Commission regulation \citep[see][for details]{eurlex}.} in the HICP, resulting in
changes of the seasonal inflation patters over time within several countries
and also between the countries. See \cite{pltm} for a detailed discussion.

\clearpage
\section{Model}\label{sec:model}

The goal of the modeling strategy, that is subsequently introduced, is the
following: detection of structural changes in the distributional properties of
the inflation rates over time, e.g., as potentially caused by interventions.
However, unlike many standard least squares approaches \citep[such as][]{bp03},
not only changes in the mean level should be addressed but also changes in
variance and potential skewness.

Hence, a likelihood-based model is adopted that can also incorporate stylized
facts for return data, such as heavy tails and skewness. Here, the seasonally
adjusted returns $y_i$ ($i = 1, \dots, n$) are assumed to come from a
generalized logistic (GL) distribution \citep[see][]{johnson} with
three-dimensional parameter $\phi_i = (\theta_i, \sigma_i, \delta_i)^\top$ for
location~$\theta_i$, scale~$\sigma_i$, and shape~$\delta_i$ at time $i$. Structural change
techniques are then employed to check whether the parameter vector $\phi_i$
remains constant over time and, if that is not the case, when and how the
parameters change. This framework allows to trace breaks in the
evolution of mean, variance and skewness of the inflation series which may
possibly be linked to underlying regime changes, e.g., in monetary policy.

For estimation of the parameters, a (quasi-)maximum likelihood framework is
adopted and the observations are assumed to be (approximately) independent,
i.e., potential autocorrelation is treated as a nuisance parameter and not
explicitly incorporated in the model. The motivation for this is to focus on the
shifts in mean, variance, and skewness -- and it turns out to be a useful model
for the data under investigation because only negligible amounts of residual
autocorrelation remain after incorporation of the structural breaks. If the
focus of the analysis were the dynamics in the autocorrelation, higher frequency
data would be required, e.g., in order to apply techniques such as those
expounded by \cite{andreou}.\footnote{For example, it would be conceivable to
obtain high-frequency data for certain years from the Billion Prices Project of
the \cite{bpp}; however, this is not pursued here.}

In the following, details are provided about (1)~the GL distribution and its properties,
(2)~how it can be estimated (under parameter stability), (3)~how parameter stability
can be assesed, and (4)~how breakpoints can be estimated in the presence of parameter
instability.

\subsection{Generalized logistic distribution}\label{sec:gld}

The logistic distribution is often used in econometrics in the
context of income and growth models due to its fatter tails compared to the 
normal distribution -- it has also been applied to the modeling of
inflation rates by \citet{batchelor}. Its generalization -- which additionally
allows for asymmetries -- has been applied to analyze risk \citep[e.g., by][]{tolikas}
but has not yet been applied -- to the best of our knowledge -- to inflation rates.
However, it will prove to be a rather simple model that fits HICP inflation rates
quite well.

The GL distribution in the parameterization used here is defined in \citet{Abberger:2002}
% The type~I GL distribution (also known as skew-logistic distribution), as defined in \citet{Abberger:2002}
%\citet{johnson},
and has the log-density function\footnote{Which is similar to the type~1 GL distribution as defined in \citet{johnson}.}
%
\begin{eqnarray}
%% f(y ~|~ \theta, \sigma, \delta) & = &
%% \frac{\frac{\delta}{\sigma} \cdot \exp\{-\frac{y - \theta}{\sigma}\}}{(1+\exp\{-\frac{y - \theta}{\sigma}\})^{(\delta+1)}} 
\ell(y ~|~ \theta, \sigma, \delta) & = & \log(\delta) - \log(\sigma) - \frac{y - \theta}{\sigma} -
                                         (\delta + 1) \cdot \log\left(1 + \exp\left\{-\frac{y - \theta}{\sigma}\right\}\right)
\label{eq:dglogis}
\end{eqnarray}
%
with location~$\theta$, scale~$\sigma > 0$ and shape~$\delta > 0$. For $\delta = 1$ the
distribution simplifies to the logistic distribution, for $\delta < 1$ or $> 1$ it is
left- or right-skewed, respectively. The corresponding first three moments are
%
\begin{eqnarray}
\mathrm{E}(y) & = & \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
\mathrm{Var}(y)  & = & \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
\mathrm{Skew}(y) & = & \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
\end{eqnarray}
%
where $\gamma(\cdot)$, $\gamma'(\cdot)$, and $\gamma''(\cdot)$ are the digamma function and its
first and second derivative, respectively.

The corresponding score function, i.e., the derivative of the log-density with respect
to the parameter vector, is given by
%
\begin{equation}
s(y ~|~ \phi) ~=~ \frac{\partial \ell(y ~|~ \phi)}{\partial \phi}  ~=~ \left(
  \begin{gathered}
    \frac{1}{\sigma} - \frac{(\delta + 1) \tilde{y}}{\sigma (1 + \tilde{y})} \\
    \left\{ \frac{1}{\sigma} - \frac{(\delta + 1) \tilde{y}}{\sigma (1 + \tilde{y})} \right\} \cdot \frac{y - \theta}{\sigma} -\frac{1}{\sigma} \\
    \frac{1}{\delta} - \log(1 + \tilde{y})
  \end{gathered} \right)
\end{equation}
%
where $\phi = (\theta, \sigma, \delta)^\top$ and
$\tilde{y} = \exp\{- (y - \theta)/\sigma\}$.

\subsection{Estimation}

Under the assumption that $y_i \sim \mathit{GL}(\theta, \sigma, \delta)$
for $i = 1, \dots, n$ independently -- i.e., are independent realizations from
a GL distribution with parameter vector $\phi = (\theta, \sigma, \delta)^\top$ --
the parameters can be estimated as usual by maximum likelihood (ML):
$\hat \phi ~=~ \argmax_\phi \sum_{i = 1}^n \ell(y_i | \phi)$. The corresponding
first order condition is $\sum_{i = 1}^n s(y_i | \hat \phi) = 0$.

To guard the inference against potential misspecification, e.g., autocorrelation or
misspecification of higher moments of the distribution, one can treat $\hat \phi$
as the quasi-maximum-likelihood (QML) estimator and adjust the inference by using
heteroskedasticity and autocorrelation consistent (HAC) covariance matrix
estimators \citep[see e.g.,][]{Andrews:1991}.

If the parameters are potentially varying over time --
$y_i \sim \mathit{GL}(\theta_i, \sigma_i, \delta_i)$ -- then $\hat \phi$ is the (Q)ML
estimator under the null hypothesis of parameter stability
%
\begin{equation} \label{eq:h0}
  H_0:~ \phi_i ~=~ \phi_0 \quad (i = 1, \dots, n)
\end{equation}
%
which should be tested against the alternative $H_1$ that at least one of the
parameters changes over time.

\subsection{Test for structural change}

The null hypothesis of parameter stability (\ref{eq:h0}) can be assessed using the empirical
scores $s(y_i | \hat \phi)$ as measures of model deviation \citep{Zeileis+Hornik:2007} from the
model fit under $H_0$. Systematic deviations over time from the full sample estimates $\hat \phi$
can then be captured by cumulative sums of the empirical scores in an empirical fluctuation
process $\mathit{efp}(t)$:
%
\begin{equation} \label{eq:efp}
\mathit{efp}(t) ~=~ \hat V^{-1/2} n^{-1/2} \sum_{i = 1}^{\lfloor nt \rfloor}
  s(y_i | \hat \phi) \quad (0 \leq t \leq 1),
\end{equation}
%
where $\hat V$ is some consistent estimator of the variance of the scores. Below,
we use the outer product of gradients $s(y_i | \hat \phi)$ to estimate $\hat V$.
Alternatively, HAC estimators could be used leading to the same qualitative
results for the data under investigation.

As usual in structural change analysis, a functional central limit theorem holds for
$\mathit{efp}(\cdot)$, which converges to a 3-dimensional Brownian
bridge: $\mathit{efp}(\cdot) ~\overset{d}{\rightarrow}~ W^0(\cdot)$. Based on
this various types of test statistics can be computed
(see \citealp{Zeileis:2005} for details). In the following, the 
sup$\mathit{LM}$ test of \cite{Andrews:1993} is employed, which performs particularly
well for single shift alternatives and if several of the three distribution
parameters change simultaneously. The test statistic with 10\% trimming is given by
%
\begin{eqnarray}
\sup_{t \in [0.1, 0.9]} \frac{\Vert \mathit{efp}(t) \Vert_2^2}{t(1-t)} \nonumber
\end{eqnarray}
%
and the appropriate $p$~values can be computed using from the corresponding
limiting distribution -- with $\mathit{efp}(t)$ replaced by $W^0(t)$.

\subsection{Breakpoint estimation}

If the null hypothesis of parameter stability (\ref{eq:h0}) is rejected, a natural
strategy is to assume that there are $B$ breakpoints with stable parameters within
each of the resulting segments. \cite{bp03} have established a rigorous inference
framework in this situation for least squares estimation which has been extended by
\cite{Zeileis+Shah+Patnaik:2010} to (Q)ML estimation. Here, we follow the same ideas
and maximize the full segmented log-likelihood
%
\begin{equation*}
  \sum_{b = 1}^{B+1} \sum_{i = \tau_b + 1}^{\tau_b} \ell(y_i ~|~ \phi^{(b)})
\end{equation*}
%
for joint estimation of the breakpoints $\tau_1, \dots, \tau_B$ and the segment-specific
GL parameters $\phi^{(b)}$ ($b = 1, \dots, B + 1$). Following the recommendations
of \cite{bp03}, a modified Bayes information criterion \citep[LWZ, proposed by][]{liu}
is employed selecting the number of breakpoints $B$.

<<analysis-results>>=
## run analysis (for seasonally adjusted data)
root_dir <- do.call("file.path", as.list(head(strsplit(getwd(), .Platform$file.sep)[[1]], -1)))
if(!file.exists(file.path(root_dir, "Analysis", "Austria.rda"))) {
setwd(file.path(root_dir, "Analysis"))
Sweave("results1s.Rnw")
setwd(file.path(root_dir, "Paper"))
}

## query result file names
fnam <- Sys.glob(file.path("..", "Analysis", "*.rda"))
fnam_omit <- c(grep("2", fnam, fixed = TRUE), grep("glmean", fnam, fixed = TRUE))
if(length(fnam_omit) > 0) fnam <- fnam[-fnam_omit]

## set up named results list
results <- vector(length = length(fnam), mode = "list")
names(results) <- sapply(strsplit(sapply(strsplit(fnam, "/", fixed = TRUE), tail, 1), ".", fixed = TRUE), head, 1)

## Exchange Rate Mechanism information
## ERM~II, ERM~I (part of european monetary system expired in 1999-01-01 gave way to ERM II
## if never entered ""right?
## http://web.archive.org/web/20080101022356/http://lexikon.meyers.de/meyers/Europ%C3%A4isches_W%C3%A4hrungssystem
erm <- yearmon(c(
  Austria = 1995,
  Belgium = 1979 + 2/12,
  CzechRepubic = NA, 
  Denmark = 1979 + 2/12,
  Estonia = NA,
  Finland = 1996 + 9/12,
  France = 1979 + 2/12,
  Germany = 1979 + 2/12,
  Greece = 1998 + 2/12,
  Hungary = NA,
  Ireland = 1979 + 2/12,
  Italy = 1979 + 2/12,
  Luxembourg = 1979 + 2/12,
  Netherlands = 1979 + 2/12,
  Poland = NA,
  Portugal = 1992 + 3/12,
  Slovakia = NA,
  Slovenia = NA,
  Spain = 1986 + 5/12,
  Sweden = NA ,
  UK = NA
))

erm2 <- yearmon(c(
  Austria = NA,
  Belgium = NA,
  CzechRepubic = NA, 
  Denmark = 1999,
  Estonia = 2004 + 5/12,
  Finland = NA,
  France = NA,
  Germany = NA,
  Greece = 1999,
  Hungary = NA,
  Ireland = NA,
  Italy = NA,
  Luxembourg = NA,
  Netherlands = NA,
  Poland = NA,
  Portugal = NA,
  Slovakia = 2005 + 10/12,
  Slovenia = 2004 + 5/12,
  Spain = NA,
  Sweden = NA ,
  UK = NA
))

eur <- yearmon(c(
  Austria = 1999,
  Belgium = 1999,
  CzechRepubic = NA, 
  Denmark = NA,
  Estonia = 2011,
  Finland = 1999,
  France = 1999,
  Germany = 1999,
  Greece = 2001,
  Hungary = NA,
  Ireland = 1999,
  Italy = 1999,
  Luxembourg = 1999,
  Netherlands = 1999,
  Poland = NA,
  Portugal = 1999,
  Slovakia = 2009,
  Slovenia = 2007,
  Spain = 1999,
  Sweden = NA ,
  UK = NA
))

for(i in seq_along(fnam)) {
  load(file.path(root_dir, "Analysis", fnam[i]))
  results[[i]] <- list(
    data = x,
    model0 = x_gf,
    efp = x_efp,
    nbreaks = x_nbreaks,
    breakpoints = x_bp,
    model = x_rf,
    erm = erm[i],
    erm2 = erm2[i],
    eur = eur[i])
}
@

\subsection{An example: Slovenia}\label{sec:slovenia}

<<slovenia>>=
x <- results$Slovenia
x$efp2 <- gefp(x$model0, fit = NULL, vcov = kernHAC)
spval <- c(sctest(x$efp, functional = supLM(0.1))$p.value,
  sctest(x$efp2, functional = supLM(0.1))$p.value)
spval <- ifelse(spval < 0.001, "$< 0.001$", format(round(spval, digits = 4), nsmall = 4))
@

To illustrate the proposed strategy, the case of Slovenia is considered. After a period of
relative instability following the independence from Yugoslavia, Slovenia was
successfull in realigning its economy and introduced a number of reforms leading
to a stable growth in recent years. The good economic performance made it
possible for Slovenia to enter the ERM~II in June~2004 and later on to introduce
the euro in January~2007. The seasonally adjusted inflation series for Slovenia encompasses
the time from \Sexpr{start(x$data)} to \Sexpr{end(x$data)} and is depicted in the left
panel of Figure~\ref{fig:slovenia1}.

\begin{figure}[t!]
<<slovenia-plots1, fig=TRUE, height=4.2, width=9.6>>=
par(mfrow = c(1, 2))
plot(x$data, main = "", ylab = "Monthly inflation rate (in percent)", xlab = "Time")
lines(x$breakpoints, breaks = x$nbreaks)
lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
if(x$nbreaks > 0) lines(confint(x$breakpoints, breaks = x$nbreaks))
plot(x$efp, functional = supLM(0.1), main = "", ylab = "LM statistics")
lines(x$breakpoints, breaks = x$nbreakss)
@
\caption{\label{fig:slovenia1} Inflation rate series for Slovenia
with breakpoint estimate (\Sexpr{breakdates(x$breakpoints)}), associated confidence interval, and fitted mean
from generalized logistic distribution (left). Corresponding sup$\mathit{LM}$ test (right)
with sequence of $\mathit{LM}$~statistics,
critical value at 5\% level (horizontal red line line), and estimated breakpoint (vertical dashed line).}

<<slovenia-plots2, fig=TRUE, height=4.2, width=9.6>>=
par(mfrow = c(1, 2))
for(i in 1:2) plot(x$model[[i]], moments = TRUE,
  xlim = c(-1.5, 2.2), ylim = c(0, 1.3),
  breaks = seq(-1.5, 2, by = 0.25),
  xlab = paste(names(x$model)[i], "\n", "Goodness-of-fit p-value: ",
    format.pval(summary(x$model[[i]])$chisq.test$p.value, digits = 4),
    sep = ""))
@
\caption{\label{fig:slovenia2} Histogramm of observed monthly inflation rates in
Slovenia for \Sexpr{names(x$model)[1]} (left) and \Sexpr{names(x$model)[1]} (right),
along with fitted generalized logistic probability density function, associated
moments, and $\chi^2$ goodness-of-fit test.}
\end{figure}

Assuming a single stable set of parmaters (\ref{eq:h0}) is  not valid, as the
sequence of $\mathit{LM}$~statistics (see right panel of Figure~\ref{fig:slovenia1}) clearly
exceeds its 5\% critical value (red horizontal line), leading to a highly significant
$p$-value \Sexpr{spval[1]}. If a quadratic spectral kernal HAC estimator \citep{Andrews:1991}
would have been used to adjust (\ref{eq:efp}), the resulting $p$-value would be somewhat
larger with \Sexpr{spval[2]} but still clearly significant.

As there is evidence for at least one break, the LWZ criterion is employed for $B = 1, 2, \dots$
breakpoints with a minimal segment size of two years per segment. It assumes its minimum for $B = 1$
associated with the breakpoint \Sexpr{breakdates(x$breakpoints, breaks = x$nbreaks)}, depcited
by a vertical dashed line in both panels of Figure~\ref{fig:slovenia1}. The resulting segmented
fitted mean (horizontal blue lines in the left panel Figure~\ref{fig:slovenia1}) shows that
in before the break Slovenia experienced very high inflation rates but was successful in reducing
inflation to a much lower level afterwards.

\clearpage
This decrease in inflation is also conveyed by Figure~\ref{fig:slovenia2} that shows the
observed histograms of inflation rates before and after the breakpoint along with the
fitted GL distribution. This highlights that along with the decrease in mean level, the variance increases
somewhat and -- probably more interestingly -- the distribution changes from being right-skewed to
slightly left-skewed. This confirms that very high inflation rates that occurred in several
months before the breakpoint could be avoided afterwards. Furthermore, Figure~\ref{fig:slovenia2}
shows that the GL distribution fits the observed data very well, both before and after the
breakpoint, with $p$~values from $\chi^2$ goodness-of-fit tests of
\Sexpr{format(round(summary(x$model[[1]])$chisq.test$p.value, digits = 3), nsmall = 3)} and
\Sexpr{format(round(summary(x$model[[2]])$chisq.test$p.value, digits = 3), nsmall = 3)}, respectively.

The estimated breakpoint can be matched very well with the timing of financial reforms
in Slovenia that severely decreased the annual growth in the M3 monetary aggregate.
These success of these actions ensured the Slovenian participation in
the ERM~II only a year later and the introduction of the euro two years later.

\clearpage
\section{Results}\label{sec:results}

The modeling strategy proposed in Section~\ref{sec:model} is applied to each of the
countries included in the data set introduced in Section~\ref{sec:data}.
Table~\ref{tab:results1} reports the estimated breakpoints along with the segment-specific
moments (mean, variance, skewness) implied by the corresponding parameter
estimates. Furthermore, entry dates for ERM/ERM~II and the euro are included as
additional information. For ease of interpretation, countries with similar patterns of
change are grouped together in Table~\ref{tab:results1} (and ordered alphabetically
within each group).

If there is only a single segment, the sup$\mathit{LM}$ test was non-significant at the 5\%
level. In significant cases, the number of breakpoints was selected via the LWZ criterion
and subsequently all parameters (breakpoints and segment-specific GL parameters) were
estimated by ML. No HAC~correction was employed as there was only negligible autocorrelation
in the empirical scores after inclusion of the breakpoints (if any).

<<results-moments>>=
## compute moments
mom <- lapply(results, function(obj) t(sapply(obj$model, moments)))
names(mom)[names(mom) == "CzechRepublic"] <- "Czech Rep."

## LaTeX formatting
format_moments <- function(i, digits = 3) {
  rval <- format(round(mom[[i]], digits = digits), nsmall = digits)
  rval <- cbind("", rownames(rval), matrix(paste("$", rval, "$", sep = ""), ncol = 3), "", "", "")
  rval[1,1] <- if(is.character(i)) i else names(mom)[i]
  rval[1,6] <- if(is.na(results[[i]]$erm)) "--" else format(results[[i]]$erm)
  rval[1,7] <- if(is.na(results[[i]]$erm2)) "--" else format(results[[i]]$erm2)
  rval[1,8] <- if(is.na(results[[i]]$eur)) "--" else format(results[[i]]$eur)
  rval <- apply(rval, 1, paste, collapse = " & ")
  rval <- paste(rval, "\\\\")
  rval[length(rval)] <- paste(rval[length(rval)], "\\hline")
  return(rval)
}
@

\begin{table}[p!]
\begin{center}
\footnotesize
\begin{tabular}{llrrrlll}
\hline
Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
\hline

\noalign{\medskip}
\multicolumn{8}{l}{\emph{No change}} \\
\hline
<<moments-table1, results=tex>>=
writeLines(unlist(lapply(c(2, 4, 8, 14), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Phase~1 convergence}} \\
\hline
<<moments-table2, results=tex>>=
writeLines(unlist(lapply(c(1, 6, 7), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Phase~2 convergence}} \\
\hline
<<moments-table3, results=tex>>=
writeLines(unlist(lapply(c(9, 12, 13, 16), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Financial crisis}} \\
\hline
<<moments-table4, results=tex>>=
writeLines(unlist(lapply(c(11, 19), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Non-euro}} \\
\hline
<<moments-table5, results=tex>>=
writeLines(unlist(lapply(c(20, 21), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Eastern countries}} \\
\hline
<<moments-table6, results=tex>>=
writeLines(unlist(lapply(c(3, 5, 10, 15, 17, 18), format_moments)))
@

\end{tabular}

\caption{\label{tab:results1} Estimated breakpoints, fitted moments of generalized logistic
distribution per segment, and ERM information for all countries under investigation.}
\end{center}
\end{table}

\subsection{The countries in detail}

For Table~\ref{tab:results1}, we grouped the countries according to either their
similar timing of the changes in inflation dynamics or according to economic similarities, 
as in the case of the former communist countries in Eastern Europe.

The first group that draws particular attention consists -- with the exception of Austria and
Luxembourg -- of the countries of the former DEM-zone. All these countries had and still have 
a very low inflation rate in terms both of mean and variance and their inflation dynamics
did not change. This does make economic sense since the ECB is build after the tradition of
the Deutsche Bundesbank and the ECB did not -- until recently -- much diverge from the 
monetary policy of the former.

Austria, Finland and France adjusted to the requirements of the EMU during the
first phase of convergence. These are also the countries with the lowest
inflation rates in both mean and variance after the adjustment. In these countries,
the result of the adjustment process was a lower inflation rate and an overall
decrease in the variance.

The next group consists of three of the PIIGS (Portugal, Ireland, Italy, 
Greece and Spain) countries and Luxembourg. Greece and
Portugal adjusted to the ERM shortly before or after its introduction.
Italy was successfull in reducing its inflation rate, whereas Luxembourg
experienced a doubling of its former -- very low -- inflation rate with the
variance rising by a factor of 14 with a strong shift in skewness towards the
left. It is possible that in the case of Luxembourg, this still reflects the
adjustment due to the treatment of sales periods, even after seasonal adjustment
of the original series.

The forth subgroup consists of two of the PIIGS  countries, namely Ireland and Spain.  
Both show a break in the summer of  2008  following the financial crisis and the severe 
problems in both countries concerning   the real estate situation. In both countries this  
had a strong deflationary impact,  with  Ireland actually showing negative inflation rates 
for an extended period of time;   in Spain, which was affected by a housing bubble of  
roughly the same size as Ireland,  the housing prices did not adjust so abruptly and  
in such a degree. This deflationary  trend, which for a time affected all countries,  
can be seen in Figure~\ref{fig:glmean},  which depicts a rolling mean estimation,  
where the mean is calculated using the first  moment of a fitted GL distribution.  
There is evidence of a decline in mean inflation,  which returns to its previous 
path roughly a year later.

\clearpage
The next group consists of two countries that decided not to be part of the
European Monetary System\footnote{If we ignore the short participating time of
both countries in the EMS.}, the United Kingdom and Sweden. Both show a clear
break in the early 90s. These can be traced back to economic crisis, the currency
crisis of the United Kingdom cumulating into the ``Black Wednesday'' in 1992 and
the banking crisis in Sweden during the early 90s.

The last group consists of the Eastern European countries: Czech Republic,
Estonia, Hungary, Poland, Slovakia and Slovenia. In almost all of these
countries, the mean and the variance of their inflation rates declined in the
later part of the 90s (Slovenia is an exception with an increase in variance
after the break). Most of them experienced a break in 1997/1998, with Poland,
Slovakia and Slovenia somewhat later. The break during the late 90s is a result
of these countries efforts to curb in inflation by decreasing the growth of the money 
supply and indicates the time when these countries actually overcame the biggest
transitionally shocks on their way towards free market economies. Slovakia is
very similar to Slovenia, as its inflation dynamics too changed roughly one year
prior to its entry into ERM~II.

\subsection{Overview over the general trend}

<<eu, echo=FALSE>>=
## impose breaks "by hand"
eu_breaks <- yearmon(c(1998 + 11/12, 2008 + 7/12))
eu_seg <- 1 + (time(eu) > eu_breaks[1]) + (time(eu) > eu_breaks[2])
library("e1071")
eu_mom_old1 <- aggregate(eu, eu_seg, function(x) c(mean(x), var(x), skewness(x)))

## better: use model-implied moments
eu_mom_old2 <- aggregate(eu, eu_seg, function(x) moments(glogisfit(x)))

## or even better: estimate breaks?
## eu_bp <- breakpoints(glogisfit(eu), h = 24)
## moments(eu_bp, breaks = 2)
## replicate results "by hand"
eu_breaks <- yearmon(c(1999 + 5/12, 2007 + 7/12))
eu_seg <- 1 + (time(eu) > eu_breaks[1]) + (time(eu) > eu_breaks[2])
eu_mom <- aggregate(eu, eu_seg, function(x) moments(glogisfit(x)))
@

The disparity of the national inflation rates within the EMU is still an issue. Many researchers, like \citet{hofmann}  find a considerable amount of inflation
differentials. \citet{cap} estimate short-run and steady-state inflation uncertainty in 12 EMU countries and find a considerable degree of heterogeneity
across EMU countries in terms of average inflation and its degree of persistence. In a paper examining structural convergence of inflation rates in EU
countries, \citet{palomba} try to answer the question if during the 1990s the inflation rate dynamics of EU countries become more similar. They find that
convergence over time of inflation dynamics was only partly observable.  In a paper studying core inflation and using an aggregated euro area inflation
rate, \citet{morana} finds three regimes (roughly 1980--1984, 1984--1993 and 1993--2000) governing the core inflation rate.

Concerning inflation differentials, Figure~\ref{fig:glmean} depicts the inflation rates and a rolling mean estimation (with a window size of 12 months),
where the mean is calculated using the first moment of a fitted GL distribution.\footnote{If more countries have the same inflation rate in a given month,
the lines are overlapped and thus get darker. We wold see one black line if all countries would show the same month-on-month increase at any one time. The
red lines depict the means.} We do see that many of these countries follow a rather similar path. One event clearly visible is the financial crises and the
short deflationary phase starting somewhere around the autumn of 2007 and lasting a few months. The general trend also suggests, that the inflation rates
converged towards some common -- albeit broad -- mean level. For the inflation series of the whole EMU, as published by the \citet{oecd}, we see
an increase in mean from
$\Sexpr{round(eu_mom[1, 1], digits = 3)}$ (\Sexpr{start(eu)}--\Sexpr{eu_breaks[1]}) to
$\Sexpr{round(eu_mom[2, 1], digits = 3)}$ (\Sexpr{eu_breaks[1] + 1/12}--\Sexpr{eu_breaks[2]}) with a slight decrease to
$\Sexpr{round(eu_mom[3, 1], digits = 3)}$ (\Sexpr{eu_breaks[2] + 1/12}--\Sexpr{end(eu)}) following the financial crises.
Whereas variance has increased considerably over time from
$\Sexpr{round(eu_mom[1, 2], digits = 3)}$ over $\Sexpr{round(eu_mom[2, 2], digits = 3)}$ to $\Sexpr{round(eu_mom[3, 2], digits = 3)}$.
Regarding skewness, we witness a movement towards negative skewness from
$\Sexpr{round(eu_mom[1, 3], digits = 3)}$ over $\Sexpr{round(eu_mom[2, 3], digits = 3)}$ to $\Sexpr{round(eu_mom[3, 3], digits = 3)}$.

<<glmean>>=
if(!file.exists(file.path(root_dir, "Analysis", "glmean.rda"))) {
   glmean <- function(x) {
    if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
  }  
  hicp_glmean <- rollapply(hicps, 12, glmean)
  setwd(file.path(root_dir, "Analysis"))
  save(hicp_glmean, file = "glmean.rda")
  setwd(file.path(root_dir, "Paper"))
} else {
  load(file.path(root_dir, "Analysis", "glmean.rda"))
}
@

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[t!]
\begin{center}
<<glmean-plot, fig=TRUE, height=5, width=7.5>>=
  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
  gray_red1 <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0))
  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1],
    xlab = "Time", ylab = "Monthly inflation rates")
  for(i in 1:ncol(hicps)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
@
\caption{\label{fig:glmean} Observed monthly inflation rates (gray) and estimated means from rolling GL fits (red)
over time for all countries.}
\end{center}
\end{figure}
\setkeys{Gin}{width=\textwidth}

<<cross-country, echo=FALSE, results=hide>>=
cross_country_moments <- function(countries, date,
  probs = c(0.25, 0.5, 0.75))
{
  ret <- mat.or.vec(9, length(date)) 
  for(jj in 1:length(date)) {
    rval <- sapply(countries, function(country) {
      nb <- results[[country]]$nbreaks 
      m <- t(sapply(results[[country]]$model, moments)) 
      bd <- if(nb < 1) { 
        end(results[[country]]$data) 
      } else {
        breakdates(results[[country]]$breakpoints, breaks = nb) 
      }
      m[sum(as.yearmon(date[jj]) > bd) + 1, ] 
    })
    v1 <- apply(rval, 1, quantile, probs = probs)
    ret[1:3, jj] <- v1[,1] 
    ret[4:6, jj] <- v1[,2] 
    ret[7:9, jj] <- v1[,3]  
  }
  return(ret)
}

format_cross_country <- function(x, digits = 3, columns = 1:NCOL(x)) {
  rval <- format(round(x, digits = digits), nsmall = digits)
  for(i in which(!(1:NCOL(rval) %in% columns))) rval[, i] <- c(rep("", 4), "\\cdots", rep("", 4))
  rval <- cbind("", "", matrix(paste("$", rval, "$", sep=""), ncol = 4))
  rval[, 1] <- c("Mean", "", "", "Variance", "", "", "Skewness", "", "")
  rval[, 2] <- rep(c("Lower", "Median", "Upper"), 3)
  rval <- apply(rval, 1, paste, collapse = " & ")
  rval <- paste(rval, "\\\\")
  rval[length(rval)] <- paste(rval[length(rval)], "\\hline")
  return(rval)
}

yrs <- c(1994, 1998, 2007, 2010)
euro <- c(1, 2, 6, 7, 8, 13, 14)
others <- c(3, 4, 5, 10, 15, 17, 18, 20, 21)
piigs <- c(9, 11, 12, 16, 19)
@

Regarding the general trend, Table~\ref{tab:results2} presents lower quartile,
median, and upper quartile (25\%, 50\%, 75\% percentiles) of the first three moments of the
inflation series in January \Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]} and \Sexpr{yrs[4]}. The countries are grouped
into the ``core'' of the eurozone (Austria, Belgium, Finland, France, Germany,
Luxembourg and Netherlands), the PIIGS countries  and all the others. For this, we took account of possible
breaks in the moments.

\begin{table}[t!]
\begin{center}
\begin{tabular}{llrrrr}
\hline
Moment & Quartile & \Sexpr{yrs[1]} & \Sexpr{yrs[2]} & \Sexpr{yrs[3]} & \Sexpr{yrs[4]}\\
\hline

\noalign{\medskip}
\multicolumn{6}{l}{\emph{Core euro}} \\
\hline
<<cross-country-table1, echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(euro, yearmon(yrs)), columns = 1:2))
@

\noalign{\medskip}
\multicolumn{6}{l}{\emph{PIIGS}} \\
\hline
<<cross-country-table2, echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(piigs, yearmon(yrs)), columns = c(1:2, 4)))
@

\noalign{\medskip}
\multicolumn{6}{l}{\emph{Other}} \\
\hline
<<cross-country-table3, echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(others, yearmon(yrs)), columns = c(1, 3)))
@

\end{tabular}
\caption{\label{tab:results2} Aggregation of estimated moments from Table~\ref{tab:results1} across countries in
January~\Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]}, and \Sexpr{yrs[4]}. Dots signal that the moments have not
changed from the previous time point, i.e., that no breakpoint was estimated in the corresponding time period.}
\end{center}
\end{table}

In the core-zone, we see a decrease in the lower quartile from \Sexpr{yrs[1]} to
\Sexpr{yrs[4]}, with no change in the median inflation rate, which is due to the
fact that there is no break after \Sexpr{yrs[2]} for these countries; however,
the interquartile range decreased by almost 20\%. The variance increased
slightly (but with a 15\% decrease in interquartile range), whereas skewness
stayed symmetric in the median, with almost no change in the interquartile
range.

The PIIGS countries decreased inflation rates as well with a strong change of skewness 
towards more negative values. The interquartile range of their inflation variance increased
considerably, which is mostly due to Spain and Ireland and their strong reaction
to the financial crisis.

The other countries were successfull in decreasing all three moments, but they
are still higher than the euro-zone countries. Most importantly, the
interquartile range for all moments decreased considerably showing a clear
convergence towards the core-zone countries.

\clearpage
\section{Conclusion}\label{sec:conclusion}

In contrast to \citet{holte}, whose simulations suggest that the standard deviation of the home 
CPI inflation rate can be substantially reduced by joining a monetary union, we fail to find a 
decrease in variance for the whole samples. However, this may be partly due to the financial crisis. 
For the countries in the core of the eurozone, the most disturbing trend seems to be the rise in 
inflation volatility in almost half of the euro countries. A very volatile inflation rate will 
likely  contribute to greater macroeconomic instability. Although \cite{grier} find no evidence 
that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA --
\cite{jaro} finds a positive correlation between the level and the standard deviation of inflation. 
This finding might be still more problematic if we take into account that on average the skewness 
parameter has decreased substantially in most of the euro-zone countries, and especially the PIIGS, 
indicating that in future higher inflation rates might be more likely. This will put pressure on the 
ECB to achieve its inflation target of 2\%. The Eastern European countries were greatly 
successfull in decreasing inflation in both mean and variance.

The results suggest that a GL distribution is appropriate for the modeling of the seasonally 
adjusted inflation rate series and that failing to account for a change in the skewness of the 
distribution might lead to wrong conclusions concerning the dynamics.

Another area for possible improvement is the data. It seems rather disturbing that adjustments 
in the computation of the HICP should lead to such considerable changes.

\clearpage
\section*{Computational details}

Our results were obtained using 
\proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} \citep{R} with the packages
\pkg{glogis}~\Sexpr{packageDescription("glogis")["Version"]},
\pkg{strucchange}~\Sexpr{packageDescription("strucchange")["Version"]} \citep{Zeileis+Leisch+Hornik:2002}, and
\pkg{fxregime}~\Sexpr{packageDescription("fxregime")["Version"]} \citep{Zeileis+Shah+Patnaik:2010},
all of which are available under the General Public License (GPL) from the
Comprehensive \proglang{R} Archive Network (\url{http://CRAN.R-project.org/}). Package \pkg{strucchange} provides
various techniques for inference in structural-change settings which are complemented
with additional maximum-likelihood-based methods by \pkg{fxregime}. Inference for the
generalized logistic model (that can be combined with \pkg{strucchange}/\pkg{fxregime})
is implemented in the new package \pkg{glogis} that also contains the data used, both
the raw monthly price indexes \citep[as provided by][]{oecd} and the seasonally adjusted
monthly inflation rates. The seasonal adjustment was performed in \pkg{X-12-ARIMA}~0.3
(\citealp{X-12-ARIMA}, see also \citealp{Findley+Monsell+Bell:1998}) through the interface
provided by \proglang{gretl}~1.9.5 (\citealp{gretl}, see also \citealp{Smith+Mixon:2006} for a review).

\bibliography{papers}

\clearpage

\begin{appendix}

\clearpage
\section{Data and sales periods}\label{sec:sales}

In addition to Germany, Portugal, Belgium, Italy, Luxembourg and Spain, there
are a number of other countries where the sales periods are regulated by law.
In France the sale periods time span is regulated by each department. In Greece there
are also two set sales-periods in winter and summer. In Italy the sales periods
are fixed each year by the Chambers of Commerce.  Portugal has two set sales
periods as well, but for two month (January and February as well as Mid-July to
Mid-September). The source for this information is \citet{konseur}. According to
\cite{pltm}, the official dates of the inclusion of these sales periods are:
January 1998 for Germany and Portugal,  January 1999 for Luxembourg and Ireland and 
January 2000 for Belgium,  while Italy and Spain introduced them in January 2001.

\clearpage
\section{Graphs}

Figures~\ref{fig:nochange}--\ref{fig:eastern} depict the data underlying the results
in Section~\ref{sec:results} along with the GL-based means from Table~\ref{tab:results1}.

<<graphs-code>>=
get_data <- function(i) {
  rval <- do.call("merge",
    lapply(i, function(j) results[[j]]$data))
  colnames(rval) <- names(results)[i]
  rval
}

get_fitted <- function(i) {
  rval <- do.call("merge",
    lapply(i, function(j) fitted(results[[j]]$breakpoints, breaks = results[[j]]$nbreaks, type = "mean")))
  colnames(rval) <- names(results)[i]
  rval
}

library("lattice")
get_xyplot <- function(i) {
  nam <- names(results)[i]
  nam[nam == "CzechRepublic"] <- "Czech Republic"
  xyplot(merge(get_data(i), get_fitted(i)),
    screen = rep(nam, 2),
    col = rep(c(1, 4), each = length(i))) 
}
@

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[p!]
\begin{center}
<<graphs-nochange, fig=TRUE, height=6.3, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(2, 4, 8, 14)))
@
\caption{\label{fig:nochange} No change.}

<<graphs-phase1, fig=TRUE, height=4.9, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(1, 6, 7)))
@
\caption{\label{fig:phase1} Phase~1 convergence.}
\end{center}
\end{figure}

\begin{figure}[p!]
\begin{center}
<<graphs-phase2, fig=TRUE, height=6.3, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(9, 12, 13, 16)))
@
\caption{\label{fig:phase2} Phase~2 convergence.}

<<graphs-crisis, fig=TRUE, height=3.5, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(11, 19)))
@
\caption{\label{fig:crisis} Financial crisis.}
\end{center}
\end{figure}

\begin{figure}[p!]
\begin{center}
<<graphs-noneuro, fig=TRUE, height=3.5, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(20, 21)))
@
\caption{\label{fig:noneuro} Non-euro.}

<<graphs-eastern, fig=TRUE, height=9.1, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(3, 5, 10, 15, 17, 18)))
@
\caption{\label{fig:eastern} Eastern countries.}
\end{center}
\end{figure}

\end{appendix}

\end{document}