### 2011-06-09 pkg/paper/Paper/draft.Rnw
draft changed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@115 b679b2f7-6539-45d0-a605-5ddcb976cde9

\documentclass[nojss]{jss}
\usepackage{amsfonts,amstext,amsmath}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}

%% need no \usepackage{Sweave}
\setkeys{Gin}{width=\textwidth}
\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide, keep.source=TRUE}

%% for internal use
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}

\title{Structural Breaks in Inflation Dynamics within the European Monetary Union}

\author{Thomas Windberger\\Universit\"at Innsbruck \And
        Achim Zeileis\\Universit\"at Innsbruck}
\Plainauthor{Thomas Windberger, Achim Zeileis}

\renewcommand{\abstractname}{Summary}
\Abstract{
	In an attempt to assess the effects of the EMU on inflation rate dynamics of its member states, this paper tests for structural breaks in the inflation dynamics of 21 European   countries. To capture   changes in the first three moments of the inflation rates, a model is adopted that makes use of the generalized logistic distribution. 
	There are   breaks in all but 4 countries.  Also, the differences between the inflation dynamics and its changes  are considerable. 
  Overall, there is a convergence towards a lower inflation rate, but with an increase in variance and a decrease in skewness.
  %% avoid "we"-constructs
  %% "adopt" better than "assume"
}

\Keywords{inflation rate, structural break, EMU, generalized logistic distribution}

\Address{
  Thomas Windberger, Achim Zeileis\\
  Department of Statistics\\
  Universit\"at Innsbruck\\
  Universit\"atsstra{\ss}e 15\\
  A-6020 Innsbruck, Austria\\
  E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
  Telephone: +43/ 512/ 507 - 7103 \\
  Fax: 43/ 512/ 507 - 2851 \\
}

\begin{document}

<<preliminaries>>=
## locale settings
if(Sys.getlocale() != "C") Sys.setlocale("LC_TIME", "English")

## packages
library("glogis")
library("fxregime")

## additional moment method
moments <- function(object, ...) UseMethod("moments")
moments.glogisfit <- function(object, ...) object$moments
moments.breakpoints.glogisfit <- function(object, breaks = NULL, ...)
  t(sapply(refit(object, breaks = breaks), "[[", "moments"))

## convenience formatting function
round3 <- function(x) format(round(x, digits = 3), nsmall = 3)
@

\section{Introduction}

\fixme{affiliation should appear on title page}
Since the beginning of the European Union, the topic of a common currency
was a controversial issue. Although the Economic and Monetary Union (EMU) is now
a fact, the discussion about the economic effects of the euro is far from being
settled. The controversial topics range from the question of whether or not the
eurozone is indeed an optimal currency area \citep[as developed in:][]{Mundell:1961}, 
all the way to the very survival of the euro in light of the budgetary problems of some of its member
states. The effects of monetary unions on a number of macroeconomic indicators,
with inflation being the most important one in this context, is in the center of
an ongoing debate. This concerns the issue of short--run and steady--state
inflation uncertainty -- in context to inflation expectations -- dealt with in
\citet{cap}, or the degree of similarity of short--run dynamic properties of the inflation rates in EU
countries, which is the topic of an investigation by \citet{palomba}.

%Any new state that wants to join the EMU is first obliged to fulfill the
%Maastricht criteria, which, besides setting rules for government debts and
%interest rates, also requires the participation in the ERM~II (Exchange Rate
%Mechanism~II) for two years, while the exchange rates towards the euro are not
%allowed to cross the nominal band.

Throughout the literature, there is still a
considerable degree of uncertainty as to what extent the introduction of the
euro, or monetary unions in general, affect the inflation rate.

Basically, the way towards the EMU basically consisted of three stages: stage~I
(1990--1994) as a phase of liberalization, stage~II (1995--1998) a phase of
convergence, and stage~III (1999--2001) the transition period, which ended with
the introduction of the euro as legal tender. A possible fourth stage is the
continuing integration of new member states, mostly former communist countries.
\readme{problem of breaks addressed} The possible influence of these different stages on the inflation rate dynamics
suggested itself for a more thorough investigation.

To inquire into possible changes in the dynamics of the inflation rate following
these stages, which may have led to considerable alterations, seems a promising task.

However, a deeper inquiry into the effects of monetary unions on the inflation rates of its member
countries is in need of a method able to
capture changes in inflation dynamics. We contribute to this field of research by
developing a new method for testing for structural breaks in the dynamics of the
inflation rates. The framework utilized in this paper allows testing for any 
possible alternations  and thus helps interpreting 
the effects of the EMU. We focus on changes in the first three moments of these 
inflation rates, which we assume to be distributed according to a generalized logistic 
distribution (GL). For this purpose, we build on previous research by \citet{Zeileis+Hornik:2007}.

This \fixme{describe method} framework allows us to assess the evolution of mean, variance, and skewness over time.
Our estimation technique enables the simultaneous estimation of breaks in these moments. 
The skewness parameter turns out to be quite important to measure alterations in the
long run inflation dynamics, as it changes considerably over the time of the
analysis. This trend would be all but invisible using a normal distribution.

%Overall, we find four periods during which structural breaks
%happened. Two of them in the 1990s (early and late), another one in the early 2000s, and the last period covers the financial crisis. There is also a 
%tendency towards a lower mean inflation rate, combined with a decreasing difference between
%individual inflation rates.

\readme{more about aggregation...} Overall, we find that the countries of the core euro--area (as defined in Section~\ref{sec:results}) did not change after the second stage and ended up with a low mean and variance. The southern euro--area member states and Ireland decreased their inflation rates almost to a level of the core countries but with a higher variance and a non--symmetric skewness. The Eastern European countries did reduce mean and variance significantly, especially the two countries that joined the euro--zone first. In the case of the Nordic non--euro countries, there is either no change at all or a clear decline in mean and variance after overcoming national crisis. Generally, inflation rates, which were very different in the beginning of the sample, are becoming more similar in terms of all three moments.

The  remainder of this paper is
structured as follows: Section~\ref{sec:data} presents the data,
Section~\ref{sec:model} presents the model and the estimation techniques used,
Section~\ref{sec:slovenia} illustrates our approach using Slovenia as an
example, Section~\ref{sec:results} presents the results, and
Section~\ref{sec:conclusion} concludes.

\clearpage
\section{Data}\label{sec:data}

<<data>>=
data("hicps", package = "glogis")
eu <- na.omit(hicps[, "EU"])
## do not include EU for figures
hicps <- hicps[, colnames(hicps) != "EU"]
@

All empirical analyses are based on seasonally adjusted inflation rates (in percent)
for \Sexpr{ncol(hicps)}~countries
(\Sexpr{gsub("UK", "United Kingdom", gsub("CzechRep", "Czech Rep", paste(colnames(hicps), collapse = ", "), fixed = TRUE), fixed = TRUE)})
in the time period from the early or mid-1990s to the end of 2010.
The original data source are seasonally unadjusted HICP series provided by the
\citet{oecd} from January~1990 (if available) to December~2010, that were transformed
to inflation rates using log-returns and subsequent seasonal adjustment
via X-12-ARIMA \citep{Findley+Monsell+Bell:1998}.

The countries in this sample can be divided into three different groups: (1)~euro
countries -- Austria, Belgium, Estonia, France, Finland, Germany, Greece,
Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia,
Spain; (2)~EU members not participating in the ERM~II -- Czech Republic,
Hungary, Poland, United Kingdom, Sweden; (3)~Denmark which stands on its own as a
member of the EU and the ERM~II, but not yet of the EMU. 
Latvia, Lithuania, Bulgaria, Romania are excluded due to data scarcity.
Cyprus, and Malta are not included since they are very small economies.

Seasonal adjustment is necessary due to the harmonization of the treatment
of sales prices\footnote{Inclusion of sales prices in the HICP was demanded by
a Commission regulation \citep[see][for details]{eurlex}.} in the HICP, resulting in
changes of the seasonal inflation patters over time within several countries
and also between the countries. See \citet{pltm} for a detailed discussion.

\clearpage
\section{Model}\label{sec:model}

The goal of the modeling strategy, that is subsequently introduced, is the
following: detection of structural changes in the distributional properties of
the inflation rates over time, e.g., as potentially caused by interventions.
However, unlike many standard least squares approaches \citep[such as][]{bp03},
not only changes in the mean level should be addressed but also changes in
variance and potential skewness.

Hence, a likelihood-based model is adopted that can also incorporate stylized
facts for return data, such as heavy tails and skewness. Here, the seasonally
adjusted returns $y_i$ ($i = 1, \dots, n$) are assumed to come from a
generalized logistic (GL) distribution \citep[see][]{johnson} with
three-dimensional parameter $\phi_i = (\theta_i, \sigma_i, \delta_i)^\top$ for
location~$\theta_i$, scale~$\sigma_i$, and shape~$\delta_i$ at time $i$. Structural change
techniques are then employed to check whether the parameter vector $\phi_i$
remains constant over time and, if that is not the case, when and how the
parameters change. This framework allows to trace breaks in the
evolution of mean, variance, and skewness of the inflation series which may
possibly be linked to underlying regime changes, e.g., in monetary policy.

For estimation of the parameters, a (quasi-)maximum likelihood framework is
adopted and the observations are assumed to be (approximately) independent,
i.e., potential autocorrelation is treated as a nuisance parameter and not
explicitly incorporated in the model. The motivation for this is to focus on the
shifts in mean, variance, and skewness -- and it turns out to be a useful model
for the data under investigation because only negligible amounts of residual
autocorrelation remain after incorporation of the structural breaks. If the
focus of the analysis were the dynamics in the autocorrelation, higher frequency
data would be required, e.g., in order to apply techniques such as those
expounded by \citet{andreou}.\footnote{For example, it would be conceivable to
obtain high-frequency data for certain years from the Billion Prices Project of
the \citet{bpp}; however, this is not pursued here.}

In the following, details are provided about (1)~the GL distribution and its properties,
(2)~how it can be estimated (under parameter stability), (3)~how parameter stability
can be assesed, and (4)~how breakpoints can be estimated in the presence of parameter
instability.

\subsection{Generalized logistic distribution}\label{sec:gld}

The logistic distribution is often used in econometrics in the
context of income and growth models due to its fatter tails compared to the 
normal distribution -- it has also been applied to the modeling of expected
inflation rates by \citet{batchelor}. Its generalization -- which additionally
allows for asymmetries -- has been applied to analyze extreme risks in the context of stock markets \citep[e.g., by][]{tolikas}
but has not yet been applied -- to the best of our knowledge -- to inflation rates.
However, it will prove to be a rather simple model that fits HICP inflation rates
quite well.

The GL distribution in the parameterization employed here is also used in \citet{Shao:2002}
% The type~I GL distribution (also known as skew-logistic distribution), as defined in \citet{Abberger:2002}
%\citet{johnson},
and has the log-density function\footnote{Which is similar to the type~1 GL distribution as defined in \citet{johnson}.}
%
\begin{eqnarray}
%% f(y ~|~ \theta, \sigma, \delta) & = &
%% \frac{\frac{\delta}{\sigma} \cdot \exp\{-\frac{y - \theta}{\sigma}\}}{(1+\exp\{-\frac{y - \theta}{\sigma}\})^{(\delta+1)}} 
\ell(y ~|~ \theta, \sigma, \delta) & = & \log(\delta) - \log(\sigma) - \frac{y - \theta}{\sigma} -
                                         (\delta + 1) \cdot \log\left(1 + \exp\left\{-\frac{y - \theta}{\sigma}\right\}\right)
\label{eq:dglogis}
\end{eqnarray}
%
with location~$\theta$, scale~$\sigma > 0$, and shape~$\delta > 0$. For $\delta = 1$ the
distribution simplifies to the logistic distribution, for $\delta < 1$ or $> 1$ it is
left- or right-skewed, respectively. The corresponding first three moments are
%
\begin{eqnarray}
\mathrm{E}(y) & = & \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
\mathrm{Var}(y)  & = & \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
\mathrm{Skew}(y) & = & \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
\end{eqnarray}
%
where $\gamma(\cdot)$, $\gamma'(\cdot)$, and $\gamma''(\cdot)$ are the digamma function and its
first and second derivative, respectively.

The corresponding score function, i.e., the derivative of the log-density with respect
to the parameter vector, is given by
%
\begin{equation}
s(y ~|~ \phi) ~=~ \frac{\partial \ell(y ~|~ \phi)}{\partial \phi}  ~=~ \left(
  \begin{gathered}
    \frac{1}{\sigma} - \frac{(\delta + 1) \tilde{y}}{\sigma (1 + \tilde{y})} \\
    \left\{ \frac{1}{\sigma} - \frac{(\delta + 1) \tilde{y}}{\sigma (1 + \tilde{y})} \right\} \cdot \frac{y - \theta}{\sigma} -\frac{1}{\sigma} \\
    \frac{1}{\delta} - \log(1 + \tilde{y})
  \end{gathered} \right)
\end{equation}
%
where $\phi = (\theta, \sigma, \delta)^\top$ and
$\tilde{y} = \exp\{- (y - \theta)/\sigma\}$.

\subsection{Estimation}

Under the assumption that $y_i \sim \mathit{GL}(\theta, \sigma, \delta)$
for $i = 1, \dots, n$ independently -- i.e., are independent realizations from
a GL distribution with parameter vector $\phi = (\theta, \sigma, \delta)^\top$ --
the parameters can be estimated as usual by maximum likelihood (ML):
$\hat \phi ~=~ \argmax_\phi \sum_{i = 1}^n \ell(y_i | \phi)$. The corresponding
first order condition is $\sum_{i = 1}^n s(y_i | \hat \phi) = 0$.

To guard the inference against potential misspecification, e.g., autocorrelation or
misspecification of higher moments of the distribution, one can treat $\hat \phi$
as the quasi-maximum-likelihood (QML) estimator and adjust the inference by using
heteroskedasticity and autocorrelation consistent (HAC) covariance matrix
estimators \citep[see e.g.,][]{Andrews:1991}.

If the parameters are potentially varying over time --
$y_i \sim \mathit{GL}(\theta_i, \sigma_i, \delta_i)$ -- then $\hat \phi$ is the (Q)ML
estimator under the null hypothesis of parameter stability
%
\begin{equation} \label{eq:h0}
  H_0:~ \phi_i ~=~ \phi_0 \quad (i = 1, \dots, n)
\end{equation}
%
which should be tested against the alternative $H_1$ that at least one of the
parameters changes over time.

\subsection{Test for structural change}

The null hypothesis of parameter stability (\ref{eq:h0}) can be assessed using the empirical
scores $s(y_i | \hat \phi)$ as measures of model deviation \citep{Zeileis+Hornik:2007} from the
model fit under $H_0$. Systematic deviations over time from the full sample estimates $\hat \phi$
can then be captured by cumulative sums of the empirical scores in an empirical fluctuation
process $\mathit{efp}(t)$:
%
\begin{equation} \label{eq:efp}
\mathit{efp}(t) ~=~ \hat V^{-1/2} n^{-1/2} \sum_{i = 1}^{\lfloor nt \rfloor}
  s(y_i ~|~ \hat \phi) \quad (0 \leq t \leq 1),
\end{equation}
%
where $\hat V$ is some consistent estimator of the variance of the scores. Below,
we use the outer product of gradients $s(y_i | \hat \phi)$ to estimate $\hat V$.
Alternatively, HAC estimators could be used leading to the same qualitative
results for the data under investigation.

As usual in structural change analysis, a functional central limit theorem holds for
$\mathit{efp}(\cdot)$, which converges to a 3-dimensional Brownian
bridge: $\mathit{efp}(\cdot) ~\overset{d}{\rightarrow}~ W^0(\cdot)$. Based on
this various types of test statistics can be computed
(see \citealp{Zeileis:2005} for details). In the following, the 
sup$\mathit{LM}$ test of \cite{Andrews:1993} is employed, which performs particularly
well for single shift alternatives and if several of the three distribution
parameters change simultaneously. The test statistic with 10\% trimming is given by
%
\begin{eqnarray}
\sup_{t \in [0.1, 0.9]} \frac{\Vert \mathit{efp}(t) \Vert_2^2}{t(1-t)} \nonumber
\end{eqnarray}
%
and the appropriate $p$~values can be computed using from the corresponding
limiting distribution -- with $\mathit{efp}(t)$ replaced by $W^0(t)$.

\subsection{Breakpoint estimation}

If the null hypothesis of parameter stability (\ref{eq:h0}) is rejected, a natural
strategy is to assume that there are $B$ breakpoints with stable parameters within
each of the resulting segments. \cite{bp03} have established a rigorous inference
framework in this situation for least squares estimation which has been extended by
\cite{Zeileis+Shah+Patnaik:2010} to (Q)ML estimation. Here, we follow the same ideas
and maximize the full segmented log-likelihood
%
\begin{equation*}
  \sum_{b = 1}^{B+1} \sum_{i = \tau_b + 1}^{\tau_b} \ell(y_i ~|~ \phi^{(b)})
\end{equation*}
%
for joint estimation of the breakpoints $\tau_1, \dots, \tau_B$ and the segment-specific
GL parameters $\phi^{(b)}$ ($b = 1, \dots, B + 1$). Following the recommendations
of \cite{bp03}, a modified Bayes information criterion \citep[LWZ, proposed by][]{liu}
is employed selecting the number of breakpoints $B$.

<<analysis-results>>=
## country names
nam <- c(colnames(hicps), "EU")

## run analysis (for seasonally adjusted data)
root_dir <- do.call("file.path", as.list(head(strsplit(getwd(), .Platform$file.sep)[[1]], -1)))
if(!all(file.exists(file.path(root_dir, "Analysis", paste(nam, "rda", sep = "."))))) {
setwd(file.path(root_dir, "Analysis"))
Sweave("results-adjusted.Rnw")
setwd(file.path(root_dir, "Paper"))
}

## set up named results list
results <- vector(length = length(nam), mode = "list")
names(results) <- nam

## Exchange Rate Mechanism information
## ERM~II, ERM~I (part of european monetary system expired in 1999-01-01 gave way to ERM II
## if never entered ""right?
## http://web.archive.org/web/20080101022356/http://lexikon.meyers.de/meyers/Europ%C3%A4isches_W%C3%A4hrungssystem
erm <- yearmon(c(
  Austria = 1995,
  Belgium = 1979 + 2/12,
  CzechRepubic = NA, 
  Denmark = 1979 + 2/12,
  Estonia = NA,
  Finland = 1996 + 9/12,
  France = 1979 + 2/12,
  Germany = 1979 + 2/12,
  Greece = 1998 + 2/12,
  Hungary = NA,
  Ireland = 1979 + 2/12,
  Italy = 1979 + 2/12,
  Luxembourg = 1979 + 2/12,
  Netherlands = 1979 + 2/12,
  Poland = NA,
  Portugal = 1992 + 3/12,
  Slovakia = NA,
  Slovenia = NA,
  Spain = 1986 + 5/12,
  Sweden = NA ,
  UK = NA,
  EU = NA
))

erm2 <- yearmon(c(
  Austria = NA,
  Belgium = NA,
  CzechRepubic = NA, 
  Denmark = 1999,
  Estonia = 2004 + 5/12,
  Finland = NA,
  France = NA,
  Germany = NA,
  Greece = 1999,
  Hungary = NA,
  Ireland = NA,
  Italy = NA,
  Luxembourg = NA,
  Netherlands = NA,
  Poland = NA,
  Portugal = NA,
  Slovakia = 2005 + 10/12,
  Slovenia = 2004 + 5/12,
  Spain = NA,
  Sweden = NA ,
  UK = NA,
  EU = NA
))

eur <- yearmon(c(
  Austria = 1999,
  Belgium = 1999,
  CzechRepubic = NA, 
  Denmark = NA,
  Estonia = 2011,
  Finland = 1999,
  France = 1999,
  Germany = 1999,
  Greece = 2001,
  Hungary = NA,
  Ireland = 1999,
  Italy = 1999,
  Luxembourg = 1999,
  Netherlands = 1999,
  Poland = NA,
  Portugal = 1999,
  Slovakia = 2009,
  Slovenia = 2007,
  Spain = 1999,
  Sweden = NA ,
  UK = NA,
  EU = NA
))

for(i in seq_along(nam)) {
  load(file.path(root_dir, "Analysis", paste(nam[i], "rda", sep = ".")))
  results[[i]] <- list(
    data = x,
    model0 = x_gf,
    efp = x_efp,
    nbreaks = x_nbreaks,
    breakpoints = x_bp,
    model = x_rf,
    erm = erm[i],
    erm2 = erm2[i],
    eur = eur[i])
}
@

\subsection{An example: Slovenia}\label{sec:slovenia}

<<slovenia>>=
x <- results$Slovenia
x$efp2 <- gefp(x$model0, fit = NULL, vcov = kernHAC)
spval <- c(sctest(x$efp, functional = supLM(0.1))$p.value,
  sctest(x$efp2, functional = supLM(0.1))$p.value)
spval <- ifelse(spval < 0.001, "$< 0.001$", format(round(spval, digits = 4), nsmall = 4))
@

To illustrate the proposed strategy, the case of Slovenia is considered. After a period of
relative instability following the independence from Yugoslavia, Slovenia was
successful in realigning its economy and introduced a number of reforms leading
to a stable growth in recent years. The good economic performance made it
possible for Slovenia to enter the ERM~II in June~2004 and later on to introduce
the euro in January~2007. The seasonally adjusted inflation series for Slovenia encompasses
the time from \Sexpr{start(x$data)} to \Sexpr{end(x$data)} and is depicted in the left
panel of Figure~\ref{fig:slovenia1}.

Assuming a single stable set of parmaters (\ref{eq:h0}) is  not valid, as the
sequence of $\mathit{LM}$~statistics (see right panel of Figure~\ref{fig:slovenia1}) clearly
exceeds its 5\% critical value (red horizontal line), leading to a highly significant
$p$-value \Sexpr{spval[1]}. If a quadratic spectral kernal HAC estimator \citep{Andrews:1991}
would have been used to adjust (\ref{eq:efp}), the resulting $p$-value would be somewhat
larger with \Sexpr{spval[2]} but still clearly significant.

As there is evidence for at least one break, the LWZ criterion is employed for $B = 1, 2, \dots$
breakpoints with a minimal segment size of two years per segment. It assumes its minimum for $B = 1$
associated with the breakpoint \Sexpr{breakdates(x$breakpoints, breaks = x$nbreaks)}, depcited
by a vertical dashed line in both panels of Figure~\ref{fig:slovenia1}. The resulting segmented
fitted mean (horizontal blue lines in the left panel Figure~\ref{fig:slovenia1}) shows that
in before the break Slovenia experienced very high inflation rates but was successful in reducing
inflation to a much lower level afterwards.

This decrease in inflation is also conveyed by Figure~\ref{fig:slovenia2} that shows the
observed histograms of inflation rates before and after the breakpoint along with the
fitted GL distribution. This highlights that along with the decrease in mean level, the variance increases
somewhat and -- probably more interestingly -- the distribution changes from being right-skewed to
slightly left-skewed. This confirms that very high inflation rates that occurred in several
months before the breakpoint could be avoided afterwards. Furthermore, Figure~\ref{fig:slovenia2}
shows that the GL distribution fits the observed data very well, both before and after the
breakpoint, with $p$~values from $\chi^2$ goodness-of-fit tests of
\Sexpr{round3(summary(x$model[[1]])$chisq.test$p.value)} and
\Sexpr{round3(summary(x$model[[2]])$chisq.test$p.value)}, respectively.

The estimated breakpoint can be matched very well with the timing of financial reforms
in Slovenia that severely decreased the annual growth in the M3 monetary aggregate.
These success of these actions ensured the Slovenian participation in
the ERM~II only a year later and the introduction of the euro two years later.

\clearpage
\section{Results}\label{sec:results}

<<cross-country, echo=FALSE, results=hide>>=
cross_country_moments <- function(countries, date,
  probs = c(0.25, 0.5, 0.75))
{
  ret <- mat.or.vec(9, length(date)) 
  for(jj in 1:length(date)) {
    rval <- sapply(countries, function(country) {
      nb <- results[[country]]$nbreaks 
      m <- t(sapply(results[[country]]$model, moments)) 
      bd <- if(nb < 1) { 
        end(results[[country]]$data) 
      } else {
        breakdates(results[[country]]$breakpoints, breaks = nb) 
      }
      m[sum(as.yearmon(date[jj]) > bd) + 1, ] 
    })
    v1 <- apply(rval, 1, quantile, probs = probs)
    ret[1:3, jj] <- v1[,1] 
    ret[4:6, jj] <- v1[,2] 
    ret[7:9, jj] <- v1[,3]  
  }
  return(ret)
}

format_cross_country <- function(x, digits = 3, columns = 1:NCOL(x)) {
  rval <- format(round(x, digits = digits), nsmall = digits)
  for(i in which(!(1:NCOL(rval) %in% columns))) rval[, i] <- c(rep("", 4), "\\cdots", rep("", 4))
  rval <- cbind("", "", matrix(paste("$", rval, "$", sep=""), ncol = 4))
  rval[, 1] <- c("Mean", "", "", "Variance", "", "", "Skewness", "", "")
  rval[, 2] <- rep(c("Lower", "Median", "Upper"), 3)
  rval <- apply(rval, 1, paste, collapse = " & ")
  rval <- paste(rval, "\\\\")
  rval[length(rval)] <- paste(rval[length(rval)], "\\hline")
  return(rval)
}

yrs <- c(1994, 1998, 2007, 2010)
euro <- c(1, 2, 6, 7, 8, 13, 14)
others <- c(3, 4, 5, 10, 15, 17, 18, 20, 21)
piigs <- c(9, 11, 12, 16, 19)
@

To investigate the effects of the EMU on national inflation rates, the
techniques expounded in Section~\ref{sec:model} are employed to
uncover changes in the dynamics of European inflation rates. The inflation
dynamics within countries from the euro-zone (but also other European
countries) is closely related with the question about the convergence
of the national inflation rates of the EMU members towards some common mean.
Many researchers provide evidence against convergence among European countries, e.g.,
\citet{hofmann} find considerable amounts of inflation differentials and
\citet{cap} describe rather strong heterogeneity among 12~EMU countries, both
in terms of average inflation and its degree of persistence.

To explore the question of convergence, we proceed in two steps:
First, in Section~\ref{sec:bpresults}, the estimation technique proposed
in Section~\ref{sec:model} is applied to each of the countries included in
the data set (see Section~\ref{sec:data}). Second, Section~\ref{sec:aggregate}
aggregates the results from each individual country to reveal different
potential convergence patterns over time within three groups of countries:
core-euro (\Sexpr{paste(colnames(hicps)[euro], collapse = ", ")}),
PIIGS  (\Sexpr{paste(colnames(hicps)[piigs], collapse = ", ")}), and the
other countries (\Sexpr{paste(colnames(hicps)[others], collapse = ", ")}).

Finally, Section~\ref{sec:fcrisis} focuses on the question of changes
in the inflation dynamics during the 2008/2009 financial crisis.

<<results-moments>>=
## compute moments
mom <- lapply(results, function(obj) t(sapply(obj$model, moments)))
names(mom)[names(mom) == "CzechRepublic"] <- "Czech Rep."

## LaTeX formatting
format_moments <- function(i, digits = 3) {
  rval <- format(round(mom[[i]], digits = digits), nsmall = digits)
  rval <- cbind("", rownames(rval), matrix(paste("$", rval, "$", sep = ""), ncol = 3), "", "", "")
  rval[1,1] <- if(is.character(i)) i else names(mom)[i]
  rval[1,6] <- if(is.na(results[[i]]$erm)) "--" else format(results[[i]]$erm)
  rval[1,7] <- if(is.na(results[[i]]$erm2)) "--" else format(results[[i]]$erm2)
  rval[1,8] <- if(is.na(results[[i]]$eur)) "--" else format(results[[i]]$eur)
  rval <- apply(rval, 1, paste, collapse = " & ")
  rval <- paste(rval, "\\\\")
  rval[length(rval)] <- paste(rval[length(rval)], "\\hline")
  return(rval)
}
@

\subsection{Inflation dynamics on a national level}\label{sec:bpresults}

As a first step towards the investigation of inflation convergence, the
structural breaks in the inflation dynamics within each countries are
investigated using the methods introduced in Section~\ref{sec:model}.
Table~\ref{tab:results1} reports the estimated breakpoints along with the
segment-specific moments (mean, variance, skewness) implied by the corresponding
parameter estimates. Furthermore, entry dates for ERM/ERM~II and the euro are
included as additional information. For ease of interpretation, countries with
similar patterns of change are grouped together and ordered alphabetically
within each group. If there is only a single segment, the sup$\mathit{LM}$ test
was non-significant at the 5\% level. In significant cases, the number of
breakpoints was selected via the LWZ criterion and subsequently all parameters
(breakpoints and segment-specific GL parameters) were estimated by ML. No
HAC~correction was employed as there was only negligible autocorrelation in the
empirical scores after inclusion of the breakpoints (if any).
In Table~\ref{tab:results1} the countries are grouped according to either a
similar timing of the changes in inflation dynamics or according to economic similarities, 
as in the case of the former communist countries in Eastern Europe.

The first group that draws particular attention consists of the former
DEM (Deutsche Mark) zone countries, with the exception of Austria and
Luxembourg. All these countries had and still have a very low inflation rate,
both in terms of mean and variance, and their inflation dynamics
did not change significantly. This is plausible from an economic point of view
since the European central bank (ECB) is modeled after the Deutsche Bundesbank and the ECB did not
-- until recently -- diverge much from the monetary policy of the former. Thus,
the EMU did have no effect on inflation dynamics since the policies in
these countries were already much in line with Germany and hence the ECB.

Austria, Finland, and France adjusted to the requirements of the EMU during the
first phase of convergence. In these countries,
the result of the adjustment process was a lower inflation rate and an overall
decrease in the variance.

The next group consists of three of the PIIGS countries and Luxembourg. Greece and
Portugal adjusted to the ERM shortly before or after its introduction.
Italy was successful in reducing its inflation rate, whereas Luxembourg
experienced a strong increase in its formerly very low inflation rate, both
in terms of mean and variances, also associated with a shift in skewness towards the
left. Apart from Luxembourg, the countries that had a structural break during the second 
phase of convergence were also successful in reducing mean inflation rates.
After the changes in this group, all countries have rather similar mean inflation
rates, closer to the levels of the previous two groups.

The fourth subgroup consists of two of the PIIGS countries, namely Ireland and Spain.  
There is a break in both countries in the summer of 2008 that captures the effects of
the financial crisis and the severe problems concerning the real estate situation.
In both countries, this had a strong deflationary impact, with deflation in Ireland
over an extended period of time. In Spain, which was affected by a housing bubble of  
roughly the same size as Ireland, the housing prices did not adjust so abruptly and  
so strongly.

The next group consists of two countries that, after a very short participation in the
ERM, decided not to be part of it and hence also not to participate in the EMU:
the United Kingdom and Sweden. Both show clear breaks in the early 1990s which can be
traced back to economic crises: the currency crisis of the United Kingdom cumulating
into the ``Black Wednesday'' in 1992 and the banking crisis in Sweden during the early
1990s.

The last group consists of the Eastern European countries: Czech Republic,
Estonia, Hungary, Poland, Slovakia, and Slovenia. In almost all of these
countries, the mean and the variance of their inflation rates declined in the
late 1990s or early 2000s. Czech Republic, Estonia and Hungary experienced a break in 1997/1998
resulting from efforts to curb in inflation by decreasing the growth of the money 
supply. The timing of the break thus indicates the timing of the biggest
transitionally shocks on the way towards free market economies. In Poland, this
transition took two more years. Slovakia and Slovenia are somewhat different: On one hand,
they started with lower mean inflation rates and on the other hand, their inflation dynamics
both changed roughly one year prior to their entry into the ERM~II.

<<glmean>>=
glmean <- function(x) {
  if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
}

if(!file.exists(file.path(root_dir, "Analysis", "glmean.rda"))) {
  hicp_glmean <- rollapply(hicps, 12, glmean)
  setwd(file.path(root_dir, "Analysis"))
  save(hicp_glmean, file = "glmean.rda")
  setwd(file.path(root_dir, "Paper"))
} else {
  load(file.path(root_dir, "Analysis", "glmean.rda"))
}
@

<<eu, echo=FALSE>>=
eu_breaks <- breakdates(results$EU$breakpoints, breaks = results$EU$nbreaks)
eu_mom <- moments(results$EU$breakpoints, breaks = results$EU$nbreaks)
@

Finally, the inflation rates for the whole euro area\footnote{This is the euro--area 17 countries series as provided by the \citet{oecd} and covers the total euro--area. This series is calculated as a weighted average of the HICPs of now seventeen euro countries and starts in January 1996.}, are considered and three different regimes are found: (1)~until 
1999 with both a low mean and variance, (2)~the third and final stage of monetary integration,
accompanied with a significant increase in the mean, (3)~a period encompassing
the overheating of the economy that culminated in the financial crisis, showing a big increase
in variance and a lower mean inflation.

\readme{perhaps skip?} One clearly visible trend is the rise in inflation volatility in some of the euro countries (Austria, Greece, Italy, Luxembourg and Spain) and the  series for the whole euro area. \citet{emerson} emphasize  the fact that a high inflation rate is also more
variable and uncertain and thus causes more relative price variability, leading
to a less efficient price mechanism. From an economic point of view, a very volatile inflation rate will 
likely  contribute to greater macroeconomic instability. In the literature, there is no consensus about the effects of higher 
inflation volatility on the inflation rate. \citet{grier} find no evidence 
that higher inflation uncertainty raises the average inflation rate -- at least in the case of the USA. However,
\cite{jaro} finds a positive correlation between the level and the standard deviation of inflation. Naturally, an
efficient price mechanism will benefit from inflation rates that do not diverge too much from the expectations.

\fixme{some concluding remarks about the EU level bridging the current gap to the next subsection}
In any case, the EU series as a whole -- which is already an aggregate -- does sum up the general trend, namely the low inflation rates during the late 90s, a higher level until the financial crisis and a lower mean inflation rate thereafter. However, this series is too highly aggregated and starts too late to be able to draw precise conclusions regarding the convergence issue.

%There is an increase in mean from
%$\Sexpr{round3(eu_mom[1, 1])}$ (\Sexpr{start(eu)}--\Sexpr{eu_breaks[1]}) to
%$\Sexpr{round3(eu_mom[2, 1])}$ (\Sexpr{eu_breaks[1] + 1/12}--\Sexpr{eu_breaks[2]}) with a slight decrease to
%$\Sexpr{round3(eu_mom[3, 1])}$ (\Sexpr{eu_breaks[2] + 1/12}--\Sexpr{end(eu)}) shortly before the financial crises.
%

\clearpage
\subsection{Convergence on an aggregate level}\label{sec:aggregate}

\fixme{intro needed}

In order to both aggregate the data to a number of sub--groups but avoid too high a level of compression, as is the case of the EU series, 
the countries are are grouped into the ``core'' of the eurozone, the PIIGS countries  and all the others. For this, we took account of possible
breaks in the moments. These groups are chosen so as to reflect the general dynamics of European inflation rates during the time period of interest.

The result of this is presented in Table~\ref{tab:results2} that shows lower quartile,
median, and upper quartile (25\%, 50\%, 75\% percentiles) of the first three moments of the
inflation series in January \Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]} and \Sexpr{yrs[4]}.

\fixme{also ``zoom out'' more, a big picture needs to be established}
The time periods are chosen according to four fundamental change periods. Namely, the early 1990s, with breaks in
Austria, Finland, France, Sweden, and UK. The late 1990s, with breaks in Greece, Italy, Luxembourg, Portugal, Czech Republic,
Estonia, and Hungary. The period after the second phase of convergence, with breaks in the early 2000s (Poland, Slovakia, and Slovenia)
and the last period consisting of two countries with the most visible effect of the financial crisis (Ireland and Spain).

In the core--zone, there is a decrease in the lower quartile from \Sexpr{yrs[1]} to
\Sexpr{yrs[2]}. However, the interquartile range decreased by almost 20\%. The variance increased
slightly, but also with a 15\% decrease in interquartile range. The skewness
stayed symmetric in the median, with almost an 8\% increase in the interquartile range.
\readme{overview}Generally, there are no dramatical changes in the core--zone and after the initial stages, none of the core--countries
had another break. This reflects the fact that their respective monetary policies were already closely connected.

The PIIGS countries significantly decreased the median inflation rates from \Sexpr{yrs[1]} to \Sexpr{yrs[4]}. 
Whereas the interquartile range was very tight in \Sexpr{yrs[1]}, it increased in \Sexpr{yrs[2]} and again in \Sexpr{yrs[4]}.
However, the median values are much smaller. The variance increased from \Sexpr{yrs[1]} to \Sexpr{yrs[2]} but did not change much 
in \Sexpr{yrs[4]}. The interquartile range did not change much too. Median skewness changed considerable 
from highly positive to negative values. 
\readme{overview, skewness effects} The decline in skewness and the narrowing of the interquartile range reflects the
fact that significant shocks to the inflation rates, which for some series were quite common in the past, have become something of a rarity. 
Still, the PIIGS countries tend to come closer towards the core--countries on behalf of the general trend of lower values and 
narrower bands.

The other countries were successful in decreasing all three moments from \Sexpr{yrs[1]} to \Sexpr{yrs[3]}. Still, their median
inflation rate is higher than those of the core countries. Also, the interquartile range of the mean decreased considerably. 
Contrary to the euro--zone and the PIIGS countries, variance decreased also, both in terms of the median value and the
interquartile range. The skewness decreased as well, as did the interquartile range of the skewness parameter.
\readme{overview} The other countries started with higher mean inflation values -- since some of them are Eastern European countries -- and 
effectually decreased the median inflation to a level much more similar to the core--zone than in \Sexpr{yrs[1]}. The bulk of changes 
in these countries took effect after \Sexpr{yrs[2]}, as a result of both an ending of the transitionary phase (from planned to free market economies) and the intended entry
into the EMU.

In context of the convergence issue, we find that, with the interquartile range of the mean of all the aggregated zones declining and getting tighter, there is  
evidence of a convergence towards some common mean inflation rate in the countries of our sample, both within the subgroups and the three groups put together.
Regarding the variance, the PIIGS countries, and the core countries as well, saw an increase over time, which is mostly due to the influence of the financial
crisis with its ``up--down--up'' turn (as depicted in Figure~\ref{fig:glmean}) leading to a higher variance.

\subsection{The influence of the financial crisis}\label{sec:fcrisis}

The results from the previous sections convey that the financial
crisis of 2008/2009 did not have a significant effect on the inflation dynamics
in most of the European countries (except Ireland and Spain).
This may be somewhat surprising and is not enitrely correct. In fact,
there is a rather clear effect but as it is very short-lived and not
persistent it is not picked up by the structural break techniques.

More specifically, the mean inflation of the euro area  from a GL fit in the 
year before the crisis (\Sexpr{yearmon(2007 + 7/12)}--\Sexpr{yearmon(2008 + 6/12)}) was
$\Sexpr{round3(glmean(window(eu, start = yearmon(2007 + 7/12), end = yearmon(2008 + 6/12))))}$
and after the crisis (\Sexpr{yearmon(2008 + 7/12)}--\Sexpr{yearmon(2009 + 6/12)}) it dropped to
$\Sexpr{round3(glmean(window(eu, start = yearmon(2008 + 7/12), end = yearmon(2009 + 6/12))))}$
before rising again in the following year  (\Sexpr{yearmon(2009 + 7/12)}--\Sexpr{yearmon(2010 + 6/12)}):
$\Sexpr{round3(glmean(window(eu, start = yearmon(2009 + 7/12), end = yearmon(2010 + 6/12))))}$.
Another way to bring out this pattern in a graphical way is
Figure~\ref{fig:glmean}. It depicts the inflation rates (in gray) and a 12-month rolling mean (in red),
calculated again as the first moment of a fitted GL distribution.
The strong increase in mean inflation during the boom years of 2007/2008
accompanied by an equally strong deflation following the burst of the bubble
is clearly visible for most countries. However, the level of mean inflation
did not change persistently, as the inflation rates seem to have returned to
their prior path, conveying the same message as the inflation series of the
euro area.

As for the structural break methods adopted in the previous sections,
this means that such rapid and non-persistent effects in the mean level
are not picked up from low frequency data, like inflation rates. Instead,
they may simply be attributed to increased variance.

Apart from bringing out more details about the short-term crisis effects,
Figure~\ref{fig:glmean} also provides an alternative view on the
long-range convergence behavior of the inflation rates studied in the
previous sections. Over the 1990s and early 2000s, there is a clear
general trend towards some common -- albeit broad -- mean level. 
There is still some visible difference in the inflation rates, but the mean
estimates are moving closer in a rather broad band. So contrary to
\citet{palomba}, who find that the degree of similarity in short-run inflation
dynamics is still weak, we find evidence of convergence within the countries in
the sample.

\clearpage
\section{Conclusion}\label{sec:conclusion}

To assess changes in the dynamics of the inflation rates of a number of European countries both within
and without the EMU, a new method for estimating structural breaks resting on the 
assumption that the inflation rates are GL distributed has been  developed.

The results suggest that a GL distribution is appropriate for the modeling of the seasonally 
adjusted inflation rate series and that failing to account for a change in the skewness of the 
distribution might lead to potentially misleading conclusions concerning the dynamics. 
The decline in size and frequency of mostly positive outliers in many inflation rate series 
would have gone unnoticed without the skewness parameter to capture these changes.

At the beginning of the sample, many countries did have very high inflation rates, especially the Eastern and Southern countries. 
For the core--zone, convergence was already achieved in \Sexpr{yrs[2]}. The PIIGS countries lowered their inflation rates considerably in preparation 
to the adoption of the euro. Over the time --  however -- many inflation rates declined considerably and are now much more in line with the inflation rates of the traditional low inflation countries of the core--euro zone. This is reflected especially well in the Eastern European countries, that lowered their high initial inflation level in the late 90s or prior to the entry into the ERM~II. They clearly benefited from the monetary discipline that is demanded for those countries who want to join the EMU.

\clearpage
\section*{Computational details}

Our results were obtained using 
\proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} \citep{R} with the packages
\pkg{glogis}~\Sexpr{packageDescription("glogis")["Version"]},
\pkg{strucchange}~\Sexpr{packageDescription("strucchange")["Version"]} \citep{Zeileis+Leisch+Hornik:2002}, and
\pkg{fxregime}~\Sexpr{packageDescription("fxregime")["Version"]} \citep{Zeileis+Shah+Patnaik:2010},
all of which are available under the General Public License (GPL) from the
Comprehensive \proglang{R} Archive Network (\url{http://CRAN.R-project.org/}). Package \pkg{strucchange} provides
various techniques for inference in structural-change settings which are complemented
with additional maximum-likelihood-based methods by \pkg{fxregime}. Inference for the
generalized logistic model (that can be combined with \pkg{strucchange}/\pkg{fxregime})
is implemented in the new package \pkg{glogis} that also contains the data used, both
the raw monthly price indexes \citep[as provided by][]{oecd} and the seasonally adjusted
monthly inflation rates. The seasonal adjustment was performed in \pkg{X-12-ARIMA}~0.3
(\citealp{X-12-ARIMA}, see also \citealp{Findley+Monsell+Bell:1998}) through the interface
provided by \proglang{gretl}~1.9.5 (\citealp{gretl}, see also \citealp{Smith+Mixon:2006} for a review).

Replication scripts for all analyses are provided within the \pkg{glogis} package: An overview
is provided in \code{demo(package = "glogis")} and the scripts can be launched easily, e.g.,
via \code{demo("Austria", package = "glogis")} etc.

\bibliography{papers}

\clearpage

\begin{appendix}

\section{Data and sales periods}\label{sec:sales}

In addition to Germany, Portugal, Belgium, Italy, Luxembourg, and Spain, there
are a number of other countries where the sales periods are regulated by law.
In France the sale periods time span is regulated by each department. In Greece there
are also two set sales-periods in winter and summer. In Italy the sales periods
are fixed each year by the Chambers of Commerce.  Portugal has two set sales
periods as well, but for two month (January and February as well as Mid-July to
Mid-September). The source for this information is \citet{konseur}. According to
\cite{pltm}, the official dates of the inclusion of these sales periods are:
January 1998 for Germany and Portugal,  January 1999 for Luxembourg and Ireland and 
January 2000 for Belgium,  while Italy and Spain introduced them in January 2001.

\clearpage
\section{Graphs}

\begin{figure}[htp!]
<<slovenia-plots1, fig=TRUE, height=4.2, width=9.6>>=
par(mfrow = c(1, 2))
plot(x$data, main = "", ylab = "Monthly inflation rate (in percent)", xlab = "Time")
lines(x$breakpoints, breaks = x$nbreaks)
lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
if(x$nbreaks > 0) lines(confint(x$breakpoints, breaks = x$nbreaks))
plot(x$efp, functional = supLM(0.1), main = "", ylab = "LM statistics")
lines(x$breakpoints, breaks = x$nbreakss)
@
\caption{\label{fig:slovenia1} Inflation rate series for Slovenia
with breakpoint estimate (\Sexpr{breakdates(x$breakpoints)}), associated confidence interval, and fitted mean
from generalized logistic distribution (left). Corresponding sup$\mathit{LM}$ test (right)
with sequence of $\mathit{LM}$~statistics,
critical value at 5\% level (horizontal red line line), and estimated breakpoint (vertical dashed line).}
\end{figure}

\clearpage

\begin{figure}[t!]
<<slovenia-plots2, fig=TRUE, height=4.2, width=9.6>>=
par(mfrow = c(1, 2))
for(i in 1:2) plot(x$model[[i]], moments = TRUE,
  xlim = c(-1.5, 2.2), ylim = c(0, 1.3),
  breaks = seq(-1.5, 2, by = 0.25),
  xlab = paste(names(x$model)[i], "\n", "Goodness-of-fit p-value: ",
    format.pval(summary(x$model[[i]])$chisq.test$p.value, digits = 4),
    sep = ""))
@
\caption{\label{fig:slovenia2} Histogramm of observed monthly inflation rates in
Slovenia for \Sexpr{names(x$model)[1]} (left) and \Sexpr{names(x$model)[1]} (right),
along with fitted generalized logistic probability density function, associated
moments, and $\chi^2$ goodness-of-fit test.}
\end{figure}

\clearpage
\begin{figure}[t!]
\begin{center}
<<glmean-plot, fig=TRUE, height=5, width=7.5>>=
  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
  gray_red1 <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0))
  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1],
    xlab = "Time", ylab = "Monthly inflation rates")
  for(i in 1:ncol(hicps)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
@
\caption{\label{fig:glmean} Observed monthly inflation rates (gray) and estimated means from
12-month rolling GL fits (red) over time for all countries.}
\end{center}
\end{figure}

\clearpage

Figures~\ref{fig:nochange}--\ref{fig:eastern} depict the data underlying the results
in Section~\ref{sec:results} along with the GL-based means from Table~\ref{tab:results1}.

<<graphs-code>>=
get_data <- function(i) {
  rval <- do.call("merge",
    lapply(i, function(j) results[[j]]$data))
  colnames(rval) <- names(results)[i]
  rval
}

get_fitted <- function(i) {
  rval <- do.call("merge",
    lapply(i, function(j) fitted(results[[j]]$breakpoints, breaks = results[[j]]$nbreaks, type = "mean")))
  colnames(rval) <- names(results)[i]
  rval
}

library("lattice")
get_xyplot <- function(i) {
  nam <- names(results)[i]
  nam[nam == "CzechRepublic"] <- "Czech Republic"
  xyplot(merge(get_data(i), get_fitted(i)),
    screen = rep(nam, 2),
    col = rep(c(1, 4), each = length(i))) 
}
@

\begin{figure}[htp!]
\begin{center}
<<graphs-nochange, fig=TRUE, height=6.3, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(2, 4, 8, 14)))
@
\caption{\label{fig:nochange} No change.}
\end{center}
\end{figure}

\clearpage

\begin{figure}[p!]
\begin{center}
<<graphs-phase1, fig=TRUE, height=4.9, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(1, 6, 7)))
@
\caption{\label{fig:phase1} Phase~1 convergence.}
\end{center}
\end{figure}

\clearpage

\begin{figure}[p!]
\begin{center}
<<graphs-phase2, fig=TRUE, height=6.3, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(9, 12, 13, 16)))
@
\caption{\label{fig:phase2} Phase~2 convergence.}
\end{center}
\end{figure}

\clearpage
\begin{figure}[p!]
\begin{center}
<<graphs-crisis, fig=TRUE, height=3.5, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(11, 19)))
@
\caption{\label{fig:crisis} Financial crisis.}
\end{center}
\end{figure}

\clearpage
\begin{figure}[p!]
\begin{center}
<<graphs-noneuro, fig=TRUE, height=3.5, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(20, 21)))
@
\caption{\label{fig:noneuro} Non-euro.}
\end{center}
\end{figure}

\clearpage
\begin{figure}[p!]
\begin{center}
<<graphs-eastern, fig=TRUE, height=9.1, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(3, 5, 10, 15, 17, 18)))
@
\caption{\label{fig:eastern} Eastern countries.}
\end{center}
\end{figure}

\clearpage
\section{Tables}

\begin{table}[t!]
\begin{center}
\footnotesize
\begin{tabular}{llrrrlll}
\hline
Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
\hline

\noalign{\medskip}
\multicolumn{8}{l}{\emph{No change}} \\
\hline
<<moments-table1, results=tex>>=
writeLines(unlist(lapply(c(2, 4, 8, 14), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Phase~1 convergence}} \\
\hline
<<moments-table2, results=tex>>=
writeLines(unlist(lapply(c(1, 6, 7), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Phase~2 convergence}} \\
\hline
<<moments-table3, results=tex>>=
writeLines(unlist(lapply(c(9, 12, 13, 16), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Financial crisis}} \\
\hline
<<moments-table4, results=tex>>=
writeLines(unlist(lapply(c(11, 19), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Non-euro}} \\
\hline
<<moments-table5, results=tex>>=
writeLines(unlist(lapply(c(20, 21), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Eastern countries}} \\
\hline
<<moments-table6, results=tex>>=
writeLines(unlist(lapply(c(3, 5, 10, 15, 17, 18), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Euro area}} \\
\hline
<<moments-table6, results=tex>>=
writeLines(unlist(lapply(22, format_moments)))
@

\end{tabular}

\caption{\label{tab:results1} Estimated breakpoints, fitted moments of generalized logistic
distribution per segment, and ERM information for all countries under investigation.}
\end{center}
\end{table}

\clearpage
\begin{table}[t!]
\begin{center}
\begin{tabular}{llrrrr}
\hline
Moment & Quartile & \Sexpr{yrs[1]} & \Sexpr{yrs[2]} & \Sexpr{yrs[3]} & \Sexpr{yrs[4]}\\
\hline

\noalign{\medskip}
\multicolumn{6}{l}{\emph{Core euro}} \\
\hline
<<cross-country-table1, echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(euro, yearmon(yrs)), columns = 1:2))
@

\noalign{\medskip}
\multicolumn{6}{l}{\emph{PIIGS}} \\
\hline
<<cross-country-table2, echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(piigs, yearmon(yrs)), columns = c(1:2, 4)))
@

\noalign{\medskip}
\multicolumn{6}{l}{\emph{Other}} \\
\hline
<<cross-country-table3, echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(others, yearmon(yrs)), columns = c(1, 3)))
@

\end{tabular}
\caption{\label{tab:results2} Aggregation of estimated moments from Table~\ref{tab:results1} across countries in
January~\Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]}, and \Sexpr{yrs[4]}. Dots signal that the moments have not
changed from the previous time point, i.e., that no breakpoint was estimated in the corresponding time period.}
\end{center}
\end{table}

\end{appendix}

\end{document}