### 2011-06-11 pkg/paper/Paper/draft.Rnw
(close to) final touches of Z
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@117 b679b2f7-6539-45d0-a605-5ddcb976cde9

\documentclass[nojss]{jss}
\usepackage{amsfonts,amstext,amsmath}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}

%% need no \usepackage{Sweave}
\setkeys{Gin}{width=\textwidth}
\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide, keep.source=TRUE}

%% for internal use
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}

\title{Structural Breaks in Inflation Dynamics within the European Monetary Union}

\author{Thomas Windberger\\Universit\"at Innsbruck \And
        Achim Zeileis\\Universit\"at Innsbruck}
%% \author{\hfill Thomas Windberger \hfill Achim Zeileis\thanks{Correspondence
%% to: Achim Zeileis, Department of Statistics,  Universit\"at Innsbruck,
%% Universit\"atsstr.~15, A-6020 Innsbruck, Austria. Tel: +43/512/507-7103. Fax: +43/512/507-2851.
%% E-mail: \email{Achim.Zeileis@R-project.org}} \hfill \hfill \\
%% {\small \it Department of Statistics, Universit\"at Innsbruck, Austria}}
\Plainauthor{Thomas Windberger, Achim Zeileis}

\renewcommand{\abstractname}{Summary}
\Abstract{
  To assess the effects of the EMU on inflation rate dynamics of
  its member states, the inflation rate series for 21~European countries
  are investigated for structural changes. To capture changes in mean,
  variance, and skewness of inflation rates, a generalized logistic
  model is adopted and complemented with structural break tests and
  breakpoint estimation techniques. These reveal considerable differences
  in the patterns of inflation dynamics and the structural changes
  therein. Overall, there is a convergence towards a lower mean inflation rate
  with reduced skewness, but accompanied by an increase in variance.
}

\Keywords{inflation rate, structural break, EMU, generalized logistic distribution}

\Address{
  Thomas Windberger, Achim Zeileis\\
  Department of Statistics\\
  Universit\"at Innsbruck\\
  Universit\"atsstra{\ss}e 15\\
  A-6020 Innsbruck, Austria\\
  E-mail: \email{Thomas.Windberger@student.uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
  Telephone: +43/12/507-7103 \\
  Fax: +43/512/507-2851 \\
}

\begin{document}

<<preliminaries>>=
## locale settings
if(Sys.getlocale() != "C") Sys.setlocale("LC_TIME", "English")

## packages
library("glogis")
library("fxregime")

## additional moment method
moments <- function(object, ...) UseMethod("moments")
moments.glogisfit <- function(object, ...) object$moments
moments.breakpoints.glogisfit <- function(object, breaks = NULL, ...)
  t(sapply(refit(object, breaks = breaks), "[[", "moments"))

## convenience formatting function
round3 <- function(x) format(round(x, digits = 3), nsmall = 3)
@

\section{Introduction}

Since the beginning of the European Union (EU), the topic of a common currency
was a controversial issue. Although the Economic and Monetary Union (EMU) is now
a fact, the discussion about the economic effects of the euro is far from being
settled. The controversial topics range from the question of whether or not the
eurozone is indeed an optimal currency area \citep[as developed in][]{Mundell:1961},
all the way to the very survival of the euro in light of
the budgetary problems of some of its member states. The effects of monetary
unions on a number of macroeconomic indicators, with inflation being of increased
importance, is at the center of an ongoing debate. This
concerns the issue of short-run and steady-state inflation uncertainty -- in relation
to inflation expectations -- as dealt with in \citet{cap}, or the degree of
similarity of short-run dynamic properties of the inflation rates in EU
countries, which is the topic of the investigation by \citet{palomba}. 
Throughout the literature, there is still a considerable degree of uncertainty
as to what extent the introduction of the euro, or monetary unions in general,
affects the inflation rate.

As the way towards the EMU essentially consisted of three stages (I:~1990--1994,
a phase of liberalization; II: 1995--1998, a phase of convergence; III: 1999--2001,
a transition period ending with the introduction of the euro as legal tender),
it seems natural to ask whether this transition was accompanied by structural changes
in the inflation dynamics in the different European countries. Furthermore, a
possible fourth stage is the continuing integration of new member states, mostly
former communist countries. Typically, changes in the mean level of the
inflation rates are of most interest but there are also reasons for investigating
changes in higher moments, such as variance and skewness. \citet{emerson} emphasize that
a high inflation rate is usually also more variable and uncertain and thus
causes more relative price variability, leading to a less stability and a less
efficient price mechanism. Furthermore, \cite{jaro} finds a positive correlation
between the level and the standard deviation of inflation.

%% From an economic point of view, a very volatile inflation rate will 
%% likely  contribute to greater macroeconomic instability. In the literature,
%% there is no consensus about the effects of higher  inflation volatility on the
%% inflation rate. \citet{grier} find no evidence that higher inflation
%% uncertainty raises the average inflation rate -- at least in the case of the
%% USA. Naturally, an efficient price mechanism will
%% benefit from inflation rates that do not diverge too much from the expectations.

%% Any new state that wants to join the EMU is first obliged to fulfill the
%% Maastricht criteria, which, besides setting rules for government debts and
%% interest rates, also requires the participation in the ERM~II (Exchange Rate
%% Mechanism~II) for two years, while the exchange rates towards the euro are not
%% allowed to cross the nominal band.

Therefore, a thorough investigation of the evolution of the inflation dynamics
within a country over time requires assessment not only of the mean level but
also higher moments. Hence, we contribute to this field of research by
developing a new method for testing for structural breaks based on a generalized
logistic (GL) distribution. This provides flexible modeling of means, variances,
and skewness, while also allowing for fat tails. Structural breaks are tested
using score-based tests \citep{Andrews:1993,Zeileis:2005} and breakpoints are estimated
by extending the least squares approach of \cite{bp03} to (quasi-)maximum likelihood
estimation of GL distributions \citep[similar in spirit to the Gaussian model
employed by][]{Zeileis+Shah+Patnaik:2010}.

The empirical results show that there are different patterns and timings of
structural breaks in the inflation dynamics of different groups of countries.
Inflation rates in the core euro area did not change significantly after
completion of the first convergence stage, yielding low mean and variance.
The Southern euro area member states (and Ireland), on the other hand, managed
to decrease the mean level of inflation in the mid-1990s, but at the price
of higher volatility. Finally, the Eastern European countries reduced both mean
and variance on their way towards joining the EMU. Generally, the inflation rates
converged from very different distributions at the beginning of the sample to
rather similar patterns at the end of the 2000s. Skewness also plays an
important role in this convergence: Some Southern and Eastern countries
have clear breaks from right-skewed to roughly symmetric inflation rates,
signalling a reduction of size and frequency of outlying high inlation rates.

The  remainder of this paper is
structured as follows: Section~\ref{sec:data} briefly presents the data,
Section~\ref{sec:model} introduces the model and the estimation techniques used,
%% Section~\ref{sec:slovenia} illustrates our approach using Slovenia as an example, 
Section~\ref{sec:results} discusses the empirical results, and
Section~\ref{sec:conclusion} concludes.

\section{Data}\label{sec:data}

<<data>>=
data("hicps", package = "glogis")
eu <- na.omit(hicps[, "EU"])
## do not include EU for figures
hicps <- hicps[, colnames(hicps) != "EU"]
@

All empirical analyses are based on seasonally adjusted inflation rates (in percent)
for \Sexpr{ncol(hicps)}~countries
(\Sexpr{gsub("UK", "United Kingdom", gsub("CzechRep", "Czech Rep", paste(colnames(hicps), collapse = ", "), fixed = TRUE), fixed = TRUE)})
in the time period from the early or mid-1990s to the end of 2010.
The original data source are seasonally unadjusted HICP series provided by the
\citet{oecd} from January~1990 (if available) to December~2010, that were transformed
to inflation rates using log-returns and subsequent seasonal adjustment
via X-12-ARIMA \citep{Findley+Monsell+Bell:1998}.

The countries in this sample can be divided into three different groups: (1)~euro
countries -- Austria, Belgium, Estonia, France, Finland, Germany, Greece,
Ireland, Italy, Luxembourg, Netherlands, Portugal, Slovakia, Slovenia,
Spain; (2)~EU members not participating in the ERM~II -- Czech Republic,
Hungary, Poland, United Kingdom, Sweden; (3)~Denmark which stands on its own as a
member of the EU and the ERM~II, but not yet of the EMU. 
Latvia, Lithuania, Bulgaria, Romania are excluded due to data scarcity.
Cyprus, and Malta are not included since they are very small economies.

Seasonal adjustment is necessary due to the harmonization of the treatment
of sales prices\footnote{Inclusion of sales prices in the HICP was demanded by
a Commission regulation \citep[see][for details]{eurlex}.} in the HICP, resulting in
changes of the seasonal inflation patters over time within several countries
and also between the countries. See \citet{pltm} for a detailed discussion.

\section{Model}\label{sec:model}

The goal of the modeling strategy, that is subsequently introduced, is the
following: detection of structural changes in the distributional properties of
the inflation rates over time, e.g., as potentially caused by interventions.
However, unlike many standard least squares approaches \citep[such as][]{bp03},
not only changes in the mean level should be addressed but also changes in
variance and potential skewness.

Hence, a likelihood-based model is adopted that can also incorporate stylized
facts for return data, such as heavy tails and skewness. Here, the seasonally
adjusted returns $y_i$ ($i = 1, \dots, n$) are assumed to come from a
generalized logistic (GL) distribution \citep[see][]{johnson} with
three-dimensional parameter $\phi_i = (\theta_i, \sigma_i, \delta_i)^\top$ for
location~$\theta_i$, scale~$\sigma_i$, and shape~$\delta_i$ at time $i$. Structural change
techniques are then employed to check whether the parameter vector $\phi_i$
remains constant over time and, if that is not the case, when and how the
parameters change. This framework allows to trace breaks in the
evolution of mean, variance, and skewness of the inflation series which may
possibly be linked to underlying regime changes, e.g., in monetary policy.

For estimation of the parameters, a (quasi-)maximum likelihood framework is
adopted and the observations are assumed to be (approximately) independent,
i.e., potential autocorrelation is treated as a nuisance parameter and not
explicitly incorporated in the model. The motivation for this is to focus on the
shifts in mean, variance, and skewness -- and it turns out to be a useful model
for the data under investigation because only negligible amounts of residual
autocorrelation remain after incorporation of the structural breaks. If the
focus of the analysis were the dynamics in the autocorrelation, higher frequency
data would be required, e.g., in order to apply techniques such as those
expounded by \citet{andreou}.\footnote{For example, it would be conceivable to
obtain high-frequency data for certain years from the Billion Prices Project of
the \citet{bpp}; however, this is not pursued here.}

In the following, details are provided about (1)~the GL distribution and its properties,
(2)~how it can be estimated (under parameter stability), (3)~how parameter stability
can be assesed, and (4)~how breakpoints can be estimated in the presence of parameter
instability.

\subsection{Generalized logistic distribution}\label{sec:gld}

The logistic distribution is often used in econometrics in the
context of income and growth models due to its fatter tails compared to the 
normal distribution -- it has also been applied to the modeling of expected
inflation rates by \citet{batchelor}. Its generalization -- which additionally
allows for asymmetries -- has been applied to analyze extreme risks in the context of stock markets \citep[e.g., by][]{tolikas}
but has not yet been applied -- to the best of our knowledge -- to inflation rates.
However, it will prove to be a rather simple model that fits HICP inflation rates
quite well.

The GL distribution employed here corresponds to the type~1 GL distribution 
from\citet{johnson}, using the parametrization and inference as in \citet{Shao:2002}. It 
has the log-density function
%
\begin{eqnarray}
%% f(y ~|~ \theta, \sigma, \delta) & = &
%% \frac{\frac{\delta}{\sigma} \cdot \exp\{-\frac{y - \theta}{\sigma}\}}{(1+\exp\{-\frac{y - \theta}{\sigma}\})^{(\delta+1)}} 
\ell(y ~|~ \theta, \sigma, \delta) & = & \log(\delta) - \log(\sigma) - \frac{y - \theta}{\sigma} -
                                         (\delta + 1) \cdot \log\left(1 + \exp\left\{-\frac{y - \theta}{\sigma}\right\}\right)
\label{eq:dglogis}
\end{eqnarray}
%
with location~$\theta$, scale~$\sigma > 0$, and shape~$\delta > 0$. For $\delta = 1$ the
distribution simplifies to the logistic distribution, for $\delta < 1$ or $> 1$ it is
left- or right-skewed, respectively. The corresponding first three moments are
%
\begin{eqnarray}
\mathrm{E}(y) & = & \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
\mathrm{Var}(y)  & = & \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
\mathrm{Skew}(y) & = & \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
\end{eqnarray}
%
where $\gamma(\cdot)$, $\gamma'(\cdot)$, and $\gamma''(\cdot)$ are the digamma function and its
first and second derivative, respectively.

The corresponding score function, i.e., the derivative of the log-density with respect
to the parameter vector, is given by
%
\begin{equation}
s(y ~|~ \phi) ~=~ \frac{\partial \ell(y ~|~ \phi)}{\partial \phi}  ~=~ \left(
  \begin{gathered}
    \frac{1}{\sigma} - \frac{(\delta + 1) \tilde{y}}{\sigma (1 + \tilde{y})} \\
    \left\{ \frac{1}{\sigma} - \frac{(\delta + 1) \tilde{y}}{\sigma (1 + \tilde{y})} \right\} \cdot \frac{y - \theta}{\sigma} -\frac{1}{\sigma} \\
    \frac{1}{\delta} - \log(1 + \tilde{y})
  \end{gathered} \right)
\end{equation}
%
where $\phi = (\theta, \sigma, \delta)^\top$ and
$\tilde{y} = \exp\{- (y - \theta)/\sigma\}$.

\subsection{Estimation}

Under the assumption that $y_i \sim \mathit{GL}(\theta, \sigma, \delta)$
for $i = 1, \dots, n$ independently -- i.e., are independent realizations from
a GL distribution with parameter vector $\phi = (\theta, \sigma, \delta)^\top$ --
the parameters can be estimated as usual by maximum likelihood (ML):
$\hat \phi ~=~ \argmax_\phi \sum_{i = 1}^n \ell(y_i | \phi)$. The corresponding
first order condition is $\sum_{i = 1}^n s(y_i | \hat \phi) = 0$.

To guard the inference against potential misspecification, e.g., autocorrelation or
misspecification of higher moments of the distribution, one can treat $\hat \phi$
as the quasi-maximum-likelihood (QML) estimator and adjust the inference by using
heteroskedasticity and autocorrelation consistent (HAC) covariance matrix
estimators \citep[see e.g.,][]{Andrews:1991}.

If the parameters are potentially varying over time --
$y_i \sim \mathit{GL}(\theta_i, \sigma_i, \delta_i)$ -- then $\hat \phi$ is the (Q)ML
estimator under the null hypothesis of parameter stability
%
\begin{equation} \label{eq:h0}
  H_0:~ \phi_i ~=~ \phi_0 \quad (i = 1, \dots, n)
\end{equation}
%
which should be tested against the alternative $H_1$ that at least one of the
parameters changes over time.

\subsection{Test for structural change}

The null hypothesis of parameter stability (\ref{eq:h0}) can be assessed using the empirical
scores $s(y_i | \hat \phi)$ as measures of model deviation \citep{Zeileis+Hornik:2007} from the
model fit under $H_0$. Systematic deviations over time from the full sample estimates $\hat \phi$
can then be captured by cumulative sums of the empirical scores in an empirical fluctuation
process $\mathit{efp}(t)$:
%
\begin{equation} \label{eq:efp}
\mathit{efp}(t) ~=~ \hat V^{-1/2} n^{-1/2} \sum_{i = 1}^{\lfloor nt \rfloor}
  s(y_i ~|~ \hat \phi) \quad (0 \leq t \leq 1),
\end{equation}
%
where $\hat V$ is some consistent estimator of the variance of the scores. Below,
we use the outer product of gradients $s(y_i | \hat \phi)$ to estimate $\hat V$.
Alternatively, HAC estimators could be used leading to the same qualitative
results for the data under investigation.

As usual in structural change analysis, a functional central limit theorem holds for
$\mathit{efp}(\cdot)$, which converges to a 3-dimensional Brownian
bridge: $\mathit{efp}(\cdot) ~\overset{d}{\rightarrow}~ W^0(\cdot)$. Based on
this various types of test statistics can be computed
(see \citealp{Zeileis:2005} for details). In the following, the 
sup$\mathit{LM}$ test of \cite{Andrews:1993} is employed, which performs particularly
well for single shift alternatives and if several of the three distribution
parameters change simultaneously. The test statistic with 10\% trimming is given by
%
\begin{eqnarray}
\sup_{t \in [0.1, 0.9]} \frac{\Vert \mathit{efp}(t) \Vert_2^2}{t(1-t)} \nonumber
\end{eqnarray}
%
and the appropriate $p$~values can be computed using from the corresponding
limiting distribution -- with $\mathit{efp}(t)$ replaced by $W^0(t)$.

\subsection{Breakpoint estimation}

If the null hypothesis of parameter stability (\ref{eq:h0}) is rejected, a natural
strategy is to assume that there are $B$ breakpoints with stable parameters within
each of the resulting segments. \cite{bp03} have established a rigorous inference
framework in this situation for least squares estimation which has been extended by
\cite{Zeileis+Shah+Patnaik:2010} to (Q)ML estimation. Here, we follow the same ideas
and maximize the full segmented log-likelihood
%
\begin{equation*}
  \sum_{b = 1}^{B+1} \sum_{i = \tau_b + 1}^{\tau_b} \ell(y_i ~|~ \phi^{(b)})
\end{equation*}
%
for joint estimation of the breakpoints $\tau_1, \dots, \tau_B$ and the segment-specific
GL parameters $\phi^{(b)}$ ($b = 1, \dots, B + 1$). Following the recommendations
of \cite{bp03}, a modified Bayes information criterion \citep[LWZ, proposed by][]{liu}
is employed selecting the number of breakpoints $B$.

<<analysis-results>>=
## country names
nam <- c(colnames(hicps), "EU")

## run analysis (for seasonally adjusted data)
root_dir <- do.call("file.path", as.list(head(strsplit(getwd(), .Platform$file.sep)[[1]], -1)))
if(!all(file.exists(file.path(root_dir, "Analysis", paste(nam, "rda", sep = "."))))) {
setwd(file.path(root_dir, "Analysis"))
Sweave("results-adjusted.Rnw")
setwd(file.path(root_dir, "Paper"))
}

## set up named results list
results <- vector(length = length(nam), mode = "list")
names(results) <- nam

## Exchange Rate Mechanism information
## ERM~II, ERM~I (part of european monetary system expired in 1999-01-01 gave way to ERM II
## if never entered ""right?
## http://web.archive.org/web/20080101022356/http://lexikon.meyers.de/meyers/Europ%C3%A4isches_W%C3%A4hrungssystem
erm <- yearmon(c(
  Austria = 1995,
  Belgium = 1979 + 2/12,
  CzechRepubic = NA, 
  Denmark = 1979 + 2/12,
  Estonia = NA,
  Finland = 1996 + 9/12,
  France = 1979 + 2/12,
  Germany = 1979 + 2/12,
  Greece = 1998 + 2/12,
  Hungary = NA,
  Ireland = 1979 + 2/12,
  Italy = 1979 + 2/12,
  Luxembourg = 1979 + 2/12,
  Netherlands = 1979 + 2/12,
  Poland = NA,
  Portugal = 1992 + 3/12,
  Slovakia = NA,
  Slovenia = NA,
  Spain = 1986 + 5/12,
  Sweden = NA ,
  UK = NA,
  EU = NA
))

erm2 <- yearmon(c(
  Austria = NA,
  Belgium = NA,
  CzechRepubic = NA, 
  Denmark = 1999,
  Estonia = 2004 + 5/12,
  Finland = NA,
  France = NA,
  Germany = NA,
  Greece = 1999,
  Hungary = NA,
  Ireland = NA,
  Italy = NA,
  Luxembourg = NA,
  Netherlands = NA,
  Poland = NA,
  Portugal = NA,
  Slovakia = 2005 + 10/12,
  Slovenia = 2004 + 5/12,
  Spain = NA,
  Sweden = NA ,
  UK = NA,
  EU = NA
))

eur <- yearmon(c(
  Austria = 1999,
  Belgium = 1999,
  CzechRepubic = NA, 
  Denmark = NA,
  Estonia = 2011,
  Finland = 1999,
  France = 1999,
  Germany = 1999,
  Greece = 2001,
  Hungary = NA,
  Ireland = 1999,
  Italy = 1999,
  Luxembourg = 1999,
  Netherlands = 1999,
  Poland = NA,
  Portugal = 1999,
  Slovakia = 2009,
  Slovenia = 2007,
  Spain = 1999,
  Sweden = NA ,
  UK = NA,
  EU = NA
))

for(i in seq_along(nam)) {
  load(file.path(root_dir, "Analysis", paste(nam[i], "rda", sep = ".")))
  results[[i]] <- list(
    data = x,
    model0 = x_gf,
    efp = x_efp,
    nbreaks = x_nbreaks,
    breakpoints = x_bp,
    model = x_rf,
    erm = erm[i],
    erm2 = erm2[i],
    eur = eur[i])
}
@

\subsection{An example: Slovenia}\label{sec:slovenia}

<<slovenia>>=
x <- results$Slovenia
x$efp2 <- gefp(x$model0, fit = NULL, vcov = kernHAC)
spval <- c(sctest(x$efp, functional = supLM(0.1))$p.value,
  sctest(x$efp2, functional = supLM(0.1))$p.value)
spval <- ifelse(spval < 0.001, "$< 0.001$", format(round(spval, digits = 4), nsmall = 4))
@

To illustrate the proposed strategy, the case of Slovenia is considered. After a period of
relative instability following the independence from Yugoslavia, Slovenia was
successful in realigning its economy and introduced a number of reforms leading
to a stable growth in recent years. The good economic performance made it
possible for Slovenia to enter the ERM~II in June~2004 and later on to introduce
the euro in January~2007. The seasonally adjusted inflation series for Slovenia encompasses
the time from \Sexpr{start(x$data)} to \Sexpr{end(x$data)} and is depicted in the left
panel of Figure~\ref{fig:slovenia1}.

\begin{figure}[t!]
<<slovenia-plots1, fig=TRUE, height=4.2, width=9.6>>=
par(mfrow = c(1, 2))
plot(x$data, main = "", ylab = "Monthly inflation rate (in percent)", xlab = "Time")
lines(x$breakpoints, breaks = x$nbreaks)
lines(fitted(x$breakpoints, breaks = x$nbreaks, type = "mean"), col = 4)
if(x$nbreaks > 0) lines(confint(x$breakpoints, breaks = x$nbreaks))
plot(x$efp, functional = supLM(0.1), main = "", ylab = "LM statistics")
lines(x$breakpoints, breaks = x$nbreakss)
@
\caption{\label{fig:slovenia1} Inflation rate series for Slovenia
with breakpoint estimate (\Sexpr{breakdates(x$breakpoints)}), associated confidence interval, and fitted mean
from generalized logistic distribution (left). Corresponding sup$\mathit{LM}$ test (right)
with sequence of $\mathit{LM}$~statistics,
critical value at 5\% level (horizontal red line line), and estimated breakpoint (vertical dashed line).}

<<slovenia-plots2, fig=TRUE, height=4.2, width=9.6>>=
par(mfrow = c(1, 2))
for(i in 1:2) plot(x$model[[i]], moments = TRUE,
  xlim = c(-1.5, 2.2), ylim = c(0, 1.3),
  breaks = seq(-1.5, 2, by = 0.25),
  xlab = paste(names(x$model)[i], "\n", "Goodness-of-fit p-value: ",
    format.pval(summary(x$model[[i]])$chisq.test$p.value, digits = 4),
    sep = ""))
@
\caption{\label{fig:slovenia2} Histogramm of observed monthly inflation rates in
Slovenia for \Sexpr{names(x$model)[1]} (left) and \Sexpr{names(x$model)[1]} (right),
along with fitted generalized logistic probability density function, associated
moments, and $\chi^2$ goodness-of-fit test.}
\end{figure}

Assuming a single stable set of parmaters (\ref{eq:h0}) is  not valid, as the
sequence of $\mathit{LM}$~statistics (see right panel of Figure~\ref{fig:slovenia1}) clearly
exceeds its 5\% critical value (red horizontal line), leading to a highly significant
$p$-value \Sexpr{spval[1]}. If a quadratic spectral kernal HAC estimator \citep{Andrews:1991}
would have been used to adjust (\ref{eq:efp}), the resulting $p$-value would be somewhat
larger with \Sexpr{spval[2]} but still clearly significant.

As there is evidence for at least one break, the LWZ criterion is employed for $B = 1, 2, \dots$
breakpoints with a minimal segment size of two years per segment. It assumes its minimum for $B = 1$
associated with the breakpoint \Sexpr{breakdates(x$breakpoints, breaks = x$nbreaks)}, depcited
by a vertical dashed line in both panels of Figure~\ref{fig:slovenia1}. The resulting segmented
fitted mean (horizontal blue lines in the left panel Figure~\ref{fig:slovenia1}) shows that
in before the break Slovenia experienced very high inflation rates but was successful in reducing
inflation to a much lower level afterwards.

This decrease in inflation is also conveyed by Figure~\ref{fig:slovenia2} that shows the
observed histograms of inflation rates before and after the breakpoint along with the
fitted GL distribution. This highlights that along with the decrease in mean level, the variance increases
somewhat and -- probably more interestingly -- the distribution changes from being right-skewed to
slightly left-skewed. This confirms that very high inflation rates that occurred in several
months before the breakpoint could be avoided afterwards. Furthermore, Figure~\ref{fig:slovenia2}
shows that the GL distribution fits the observed data very well, both before and after the
breakpoint, with $p$~values from $\chi^2$ goodness-of-fit tests of
\Sexpr{round3(summary(x$model[[1]])$chisq.test$p.value)} and
\Sexpr{round3(summary(x$model[[2]])$chisq.test$p.value)}, respectively.

The estimated breakpoint can be matched very well with the timing of financial reforms
in Slovenia that severely decreased the annual growth in the M3 monetary aggregate.
These success of these actions ensured the Slovenian participation in
the ERM~II only a year later and the introduction of the euro two years later.

\section{Results}\label{sec:results}

<<cross-country, echo=FALSE, results=hide>>=
cross_country_moments <- function(countries, date,
  probs = c(0.25, 0.5, 0.75))
{
  ret <- mat.or.vec(9, length(date)) 
  for(jj in 1:length(date)) {
    rval <- sapply(countries, function(country) {
      nb <- results[[country]]$nbreaks 
      m <- t(sapply(results[[country]]$model, moments)) 
      bd <- if(nb < 1) { 
        end(results[[country]]$data) 
      } else {
        breakdates(results[[country]]$breakpoints, breaks = nb) 
      }
      m[sum(as.yearmon(date[jj]) > bd) + 1, ] 
    })
    v1 <- apply(rval, 1, quantile, probs = probs)
    ret[1:3, jj] <- v1[,1] 
    ret[4:6, jj] <- v1[,2] 
    ret[7:9, jj] <- v1[,3]  
  }
  return(ret)
}

format_cross_country <- function(x, digits = 3, columns = 1:NCOL(x)) {
  rval <- format(round(x, digits = digits), nsmall = digits)
  for(i in which(!(1:NCOL(rval) %in% columns))) rval[, i] <- c(rep("", 4), "\\cdots", rep("", 4))
  rval <- cbind("", "", matrix(paste("$", rval, "$", sep=""), ncol = 4))
  rval[, 1] <- c("Mean", "", "", "Variance", "", "", "Skewness", "", "")
  rval[, 2] <- rep(c("Lower", "Median", "Upper"), 3)
  rval <- apply(rval, 1, paste, collapse = " & ")
  rval <- paste(rval, "\\\\")
  rval[length(rval)] <- paste(rval[length(rval)], "\\hline")
  return(rval)
}

yrs <- c(1994, 1998, 2007, 2010)
euro <- c(1, 2, 6, 7, 8, 13, 14)
others <- c(3, 4, 5, 10, 15, 17, 18, 20, 21)
piigs <- c(9, 11, 12, 16, 19)
@

To investigate the effects of the EMU on national inflation rates, the
techniques expounded in Section~\ref{sec:model} are employed to
uncover changes in the dynamics of European inflation rates. The inflation
dynamics within countries from the euro-zone (but also other European
countries) is closely related with the question about the convergence
of the national inflation rates of the EMU members towards some common mean.
Many researchers provide evidence against convergence among European countries, e.g.,
\citet{hofmann} find considerable amounts of inflation differentials and
\citet{cap} describe rather strong heterogeneity among 12~EMU countries, both
in terms of average inflation and its degree of persistence.

To explore the question of convergence, we proceed in two steps:
First, in Section~\ref{sec:bpresults}, the estimation technique proposed
in Section~\ref{sec:model} is applied to each of the countries included in
the data set (see Section~\ref{sec:data}). Second, Section~\ref{sec:aggregate}
aggregates the results from each individual country to reveal different
potential convergence patterns over time within three groups of countries:
core-euro (\Sexpr{paste(colnames(hicps)[euro], collapse = ", ")}),
PIIGS  (\Sexpr{paste(colnames(hicps)[piigs[c(4, 2, 3, 1, 5)]], collapse = ", ")}), and the
other countries (\Sexpr{paste(gsub("CzechRepublic", "Czech Republic", colnames(hicps))[others], collapse = ", ")}).

Finally, Section~\ref{sec:fcrisis} focuses on the question of changes
in the inflation dynamics during the 2008/2009 financial crisis.

<<results-moments>>=
## compute moments
mom <- lapply(results, function(obj) t(sapply(obj$model, moments)))
names(mom)[names(mom) == "CzechRepublic"] <- "Czech Rep."

## LaTeX formatting
format_moments <- function(i, digits = 3) {
  rval <- format(round(mom[[i]], digits = digits), nsmall = digits)
  rval <- cbind("", rownames(rval), matrix(paste("$", rval, "$", sep = ""), ncol = 3), "", "", "")
  rval[1,1] <- if(is.character(i)) i else names(mom)[i]
  rval[1,6] <- if(is.na(results[[i]]$erm)) "--" else format(results[[i]]$erm)
  rval[1,7] <- if(is.na(results[[i]]$erm2)) "--" else format(results[[i]]$erm2)
  rval[1,8] <- if(is.na(results[[i]]$eur)) "--" else format(results[[i]]$eur)
  rval <- apply(rval, 1, paste, collapse = " & ")
  rval <- paste(rval, "\\\\")
  rval[length(rval)] <- paste(rval[length(rval)], "\\hline")
  return(rval)
}
@

\begin{table}[p!]
\begin{center}
\footnotesize
\begin{tabular}{llrrrlll}
\hline
Country & Segment & Mean & Variance & Skewness & ERM & ERM~II & Euro \\
\hline

\noalign{\medskip}
\multicolumn{8}{l}{\emph{No change}} \\
\hline
<<moments-table1, results=tex>>=
writeLines(unlist(lapply(c(2, 4, 8, 14), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Phase~1 convergence}} \\
\hline
<<moments-table2, results=tex>>=
writeLines(unlist(lapply(c(1, 6, 7), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Phase~2 convergence}} \\
\hline
<<moments-table3, results=tex>>=
writeLines(unlist(lapply(c(9, 12, 13, 16), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Financial crisis}} \\
\hline
<<moments-table4, results=tex>>=
writeLines(unlist(lapply(c(11, 19), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Non-euro}} \\
\hline
<<moments-table5, results=tex>>=
writeLines(unlist(lapply(c(20, 21), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Eastern countries}} \\
\hline
<<moments-table6, results=tex>>=
writeLines(unlist(lapply(c(3, 5, 10, 15, 17, 18), format_moments)))
@

\noalign{\medskip}
\multicolumn{8}{l}{\emph{Euro area}} \\
\hline
<<moments-table6, results=tex>>=
writeLines(unlist(lapply(22, format_moments)))
@

\end{tabular}

\caption{\label{tab:results1} Estimated breakpoints, fitted moments of generalized logistic
distribution per segment, and ERM information for all countries under investigation.}
\end{center}
\end{table}

\subsection{Inflation dynamics on a national level}\label{sec:bpresults}

As a first step towards the investigation of inflation convergence, the
structural breaks in the inflation dynamics within each countries are
investigated using the methods introduced in Section~\ref{sec:model}.
Table~\ref{tab:results1} reports the estimated breakpoints along with the
segment-specific moments (mean, variance, skewness) implied by the corresponding
parameter estimates. Furthermore, entry dates for ERM/ERM~II and the euro are
included as additional information. For ease of interpretation, countries with
similar patterns of change are grouped together and ordered alphabetically
within each group. If there is only a single segment, the sup$\mathit{LM}$ test
was non-significant at the 5\% level. In significant cases, the number of
breakpoints was selected via the LWZ criterion and subsequently all parameters
(breakpoints and segment-specific GL parameters) were estimated by ML. No
HAC~correction was employed as there was only negligible autocorrelation in the
empirical scores after inclusion of the breakpoints (if any).
In Table~\ref{tab:results1} the countries are grouped according to either a
similar timing of the changes in inflation dynamics or according to economic similarities, 
as in the case of the former communist countries in Eastern Europe.

The first group that draws particular attention consists of the former
DEM (Deutsche Mark) zone countries, with the exception of Austria and
Luxembourg. All these countries had and still have a very low inflation rate,
both in terms of mean and variance, and their inflation dynamics
did not change significantly. This is plausible from an economic point of view
since the European central bank (ECB) is modeled after the Deutsche Bundesbank and the ECB did not
-- until recently -- diverge much from the monetary policy of the former. Thus,
the EMU did have no effect on inflation dynamics since the policies in
these countries were already much in line with Germany and hence the ECB.

Austria, Finland, and France adjusted to the requirements of the EMU during the
first phase of convergence. In these countries,
the result of the adjustment process was a lower inflation rate and an overall
decrease in the variance.

The next group consists of three of the PIIGS countries and Luxembourg. Greece and
Portugal adjusted to the ERM shortly before or after its introduction.
Italy was successful in reducing its inflation rate, whereas Luxembourg
experienced a strong increase in its formerly very low inflation rate, both
in terms of mean and variances, also associated with a shift in skewness towards the
left. Apart from Luxembourg, the countries that had a structural break during the second 
phase of convergence were also successful in reducing mean inflation rates.
After the changes in this group, all countries have rather similar mean inflation
rates, closer to the levels of the previous two groups.

The fourth subgroup consists of two of the PIIGS countries, namely Ireland and Spain.  
There is a break in both countries in the summer of 2008 that captures the effects of
the financial crisis and the severe problems concerning the real estate situation.
In both countries, this had a strong deflationary impact, with deflation in Ireland
over an extended period of time. In Spain, which was affected by a housing bubble of  
roughly the same size as Ireland, the housing prices did not adjust so abruptly and  
so strongly.

The next group consists of two countries that, after a very short participation in the
ERM, decided not to be part of it and hence also not to participate in the EMU:
the United Kingdom and Sweden. Both show clear breaks in the early 1990s which can be
traced back to economic crises: the currency crisis of the United Kingdom cumulating
into the ``Black Wednesday'' in 1992 and the banking crisis in Sweden during the early
1990s.

The last group consists of the Eastern European countries: Czech Republic,
Estonia, Hungary, Poland, Slovakia, and Slovenia. In almost all of these
countries, the mean and the variance of their inflation rates declined in the
late 1990s or early 2000s. Czech Republic, Estonia and Hungary experienced a break in 1997/1998
resulting from efforts to curb in inflation by decreasing the growth of the money 
supply. The timing of the break thus indicates the timing of the biggest
transitionally shocks on the way towards free market economies. In Poland, this
transition took two more years. Slovakia and Slovenia are somewhat different: On one hand,
they started with lower mean inflation rates and on the other hand, their inflation dynamics
both changed roughly one year prior to their entry into the ERM~II.

<<glmean>>=
glmean <- function(x) {
  if(any(is.na(x))) NA else glogisfit(x, hessian = FALSE)$moments[1]
}

if(!file.exists(file.path(root_dir, "Analysis", "glmean.rda"))) {
  hicp_glmean <- rollapply(hicps, 12, glmean)
  setwd(file.path(root_dir, "Analysis"))
  save(hicp_glmean, file = "glmean.rda")
  setwd(file.path(root_dir, "Paper"))
} else {
  load(file.path(root_dir, "Analysis", "glmean.rda"))
}
@

<<eu, echo=FALSE>>=
eu_breaks <- breakdates(results$EU$breakpoints, breaks = results$EU$nbreaks)
eu_mom <- moments(results$EU$breakpoints, breaks = results$EU$nbreaks)
@

Finally, the inflation rates for the whole euro area\footnote{This is the
euro area 17~countries series as provided by the \citet{oecd},
starting in \Sexpr{start(eu) - 1/12}. It is calculated as a weighted average of the HICPs
of now seventeen euro countries.} are considered and
three different regimes are found: (1)~until 1999 with both a low mean and
variance, (2)~the third and final stage of monetary integration, accompanied
by a significant increase in the mean, (3)~a period encompassing the
overheating of the economy and its culmination in the financial crisis, showing a
large increase in variance and a lower mean inflation.

\subsection{Convergence on an aggregate level}\label{sec:aggregate}

\begin{table}[t!]
\begin{center}
\begin{tabular}{llrrrr}
\hline
Moment & Quartile & \Sexpr{yrs[1]} & \Sexpr{yrs[2]} & \Sexpr{yrs[3]} & \Sexpr{yrs[4]}\\
\hline

\noalign{\medskip}
\multicolumn{6}{l}{\emph{Core euro}} \\
\hline
<<cross-country-table1, echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(euro, yearmon(yrs)), columns = 1:2))
@

\noalign{\medskip}
\multicolumn{6}{l}{\emph{PIIGS}} \\
\hline
<<cross-country-table2, echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(piigs, yearmon(yrs)), columns = c(1:2, 4)))
@

\noalign{\medskip}
\multicolumn{6}{l}{\emph{Other}} \\
\hline
<<cross-country-table3, echo=FALSE, results=tex>>=
writeLines(format_cross_country(cross_country_moments(others, yearmon(yrs)), columns = c(1, 3)))
@

\end{tabular}
\caption{\label{tab:results2} Aggregation of estimated moments from Table~\ref{tab:results1} across countries in
Jan~\Sexpr{yrs[1]}, \Sexpr{yrs[2]}, \Sexpr{yrs[3]}, and \Sexpr{yrs[4]}. Dots signal that the moments have not
changed from the previous time point, i.e., that no breakpoint was estimated in the corresponding time period.}
\end{center}
\end{table}

Given the results above for the individual countries, a natural question is
whether there is convergence on an aggregate level and, if so, which pattern it
follows. Hence, we aggregate the results from Table~\ref{tab:results1} for the
individual countries into three groups and at four time points to reflect the
general dynamics of European inflation rates during the time period of interest.
The groups have been selected based on prior expectations (i.e., not based on
the breakpoints estimation results) as the ``core'' of the euro zone
(\Sexpr{paste(colnames(hicps)[euro], collapse = ", ")}), the PIIGS countries,
and the rest. Similarly, the time points have been chosen as to fall ``between''
times where transitions in one or more groups of countries could be expected:
Jan~\Sexpr{yrs[1]} during stage~I on the way towards the EMU, Jan~\Sexpr{yrs[2]}
during stage~II, Jan~\Sexpr{yrs[3]} after transition into the EMU but before the
financial crisis, and Jan~\Sexpr{yrs[4]} after the crisis. The result of these
considerations is presented in Table~\ref{tab:results2} which extracts the three
moments for each country from Table~\ref{tab:results1} at the indicated time
points and aggregates them within the group. Lower quartile, median, and upper
quartile (25\%, 50\%, 75\% percentiles) are provided to assess both changes in
location and spread across countries.

For the core euro countries, there are only changes from stage~I to stage~II
because after \Sexpr{yrs[2]} no breaks have been detected. The changes in this
first phase did not affect mean inflation much, while variance increased
slightly (in both quartiles) and skewness somewhat decreased (from slightly
right- to slightly left-skewed). The heterogeneity (assessed by the
interquartile range) decreased for all three moments of the inflation rates.
This reflects that the monetary policies of the corresponding countries were
already closely connected prior to stage~I and now even moved somewhat closer
together.

For the PIIGS countries, changes occured from \Sexpr{yrs[1]} to
\Sexpr{yrs[2]} and then again during the crisis, before \Sexpr{yrs[4]}.
In the first phase, mean inflation was almost halved, accompanied by a
doubling of variance and a change from clearly right-skewed to approximately
symmetric. The latter reflects that the occurence of shocks with high inflation
rates have been successfully reduced. Moreover, the changes in the PIIGS
countries between \Sexpr{yrs[3]} and \Sexpr{yrs[4]} lead to a further
decrease in mean inflation and associated skewness which are both caused by
several extremely low inflation (or even deflation) rates during the time
following the crisis.

The other countries do not exhibit any significant breakpoints during the
mid-1990s, but many changes occur in the transition period from \Sexpr{yrs[2]}
to \Sexpr{yrs[3]}, encompassing both the completion of transformation from
planned to free market economies and the intended entry into the EMU.
During this time, the countries were successful in
decreasing all three moments of their inflation distributions while
simultaneously decreasing the heterogeneity (as captured by the corresponding
interquartile ranges) across countries. Before the transition all three
moments were extremely large compared to the other two groups, while afterwards
all moments are much closer to those of the other groups. In Jan~\Sexpr{yrs[3]},
mean inflation is in between the levels for the core euro and PIIGS countres,
variance is still slightly higher, and skewness is not too far from symmetry.

In context of the convergence issue, we find that up to \Sexpr{yrs[3]}
the inflation dynamics of all three groups of countries become much
more similar, both within and across groups. The interquartile ranges
of the means decline and become tighter for all groups and there appears to
be convergence towards some common mean inflation rate in the
countries of our sample. Variances also move more closely together, corresponding
to an increase for the core euro and PIIGS countries but a clear decrease
for the others. Right-skewness (and thus large outlying inflation rates)
has been reduced for all three groups of countries but to different extents.
Thus, while ending in similar inflation dynamics, the timing and patterns of
changes in this convergence process are rather different across
the three groups, depending on the different starting situations of the
countries.

\subsection{The influence of the financial crisis}\label{sec:fcrisis}

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[t!]
\begin{center}
<<glmean-plot, fig=TRUE, height=5, width=7.5>>=
  gray_red <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0), alpha = 0.3)
  gray_red1 <- rgb(c(0.2, 0.8), c(0.2, 0), c(0.2, 0))
  plot(hicps, plot.type = "single", lwd = 1.5, col = gray_red[1],
    xlab = "Time", ylab = "Monthly inflation rates")
  for(i in 1:ncol(hicps)) lines(hicp_glmean[,i], lwd = 1.5, col = gray_red[2])
@
\caption{\label{fig:glmean} Observed monthly inflation rates (gray) and estimated means from
12-month rolling GL fits (red) over time for all countries.}
\end{center}
\end{figure}
\setkeys{Gin}{width=\textwidth}

The results from the previous sections convey that the financial
crisis of 2008/2009 did not have a significant effect on the inflation dynamics
in most of the European countries (except Ireland and Spain).
This may be somewhat surprising and is not enitrely correct. In fact,
there is a rather clear effect but as it is very short-lived and not
persistent it is not picked up by the structural break techniques.

More specifically, the mean inflation of the euro area  from a GL fit in the 
year before the crisis (\Sexpr{yearmon(2007 + 7/12)}--\Sexpr{yearmon(2008 + 6/12)}) was
$\Sexpr{round3(glmean(window(eu, start = yearmon(2007 + 7/12), end = yearmon(2008 + 6/12))))}$
and after the crisis (\Sexpr{yearmon(2008 + 7/12)}--\Sexpr{yearmon(2009 + 6/12)}) it dropped to
$\Sexpr{round3(glmean(window(eu, start = yearmon(2008 + 7/12), end = yearmon(2009 + 6/12))))}$
before rising again in the following year  (\Sexpr{yearmon(2009 + 7/12)}--\Sexpr{yearmon(2010 + 6/12)}):
$\Sexpr{round3(glmean(window(eu, start = yearmon(2009 + 7/12), end = yearmon(2010 + 6/12))))}$.
Another way to bring out this pattern in a graphical way is
Figure~\ref{fig:glmean}. It depicts the inflation rates (in gray) and a 12-month rolling mean (in red),
calculated again as the first moment of a fitted GL distribution.
The strong increase in mean inflation during the boom years of 2007/2008
accompanied by an equally strong deflation following the burst of the bubble
is clearly visible for most countries. However, the level of mean inflation
did not change persistently, as the inflation rates seem to have returned to
their prior path, conveying the same message as the inflation series of the
euro area.

As for the structural break methods adopted in the previous sections,
this means that such rapid and non-persistent effects in the mean level
are not picked up from low frequency data, like inflation rates. Instead,
they may simply be attributed to increased variance.

Apart from bringing out more details about the short-term crisis effects,
Figure~\ref{fig:glmean} also provides an alternative view on the
long-range convergence behavior of the inflation rates studied in the
previous sections. Over the 1990s and early 2000s, there is a clear
general trend towards some common -- albeit broad -- mean level. 
There is still some visible difference in the inflation rates, but the mean
estimates are moving closer in a rather broad band. So contrary to
\citet{palomba}, who find that the degree of similarity in short-run inflation
dynamics is still weak, we find evidence of convergence within the countries in
the sample.

\section{Conclusion}\label{sec:conclusion}

To assess changes in the dynamics of the inflation rates of a number of European countries
-- both within and outside the EMU -- a new method for estimating structural breaks in
inflation rates is proposed. The method allows to test for changes in mean, variance,
and skewness of inflation rate series by adopting a (quasi-)likelihood model based on the generalized
logistic (GL) distribution. Thus, unlike normally distributed models, this approach
can deal with fat tails and (potentially changing) skewness.

The empirical results suggest that a GL distribution is appropriate for modeling seasonally 
adjusted inflation rate series and is able to uncover important changes in all three moments.
Towards the end of the 2000s, before the financial crisis, our results broadly confirm
convergence hypotheses as the first three moments from all countries in our sample
move closer together.
Especially the changes in skewness -- caused by a reduction of size and frequency of outlying high inflation
rates in some countries -- might have gone unnoticed with traditional structural change techniques,
leading to potentially misleading conclusions concerning the dynamics.

At the beginning of the sample, many PIIGS and Eastern countries did have very high inflation
rates, while for the core euro countries, convergence was already achieved in the mid-1990s.
The PIIGS countries lowered their inflation rates considerably in preparation to the adoption
of the euro. The Eastern European countries lowered their high initial inflation level in the
late 1990s or early 2000s, prior to the entry into the ERM~II. They clearly benefited from the
monetary discipline that is demanded for those countries who want to join the EMU.

\section*{Computational details}

Our results were obtained using 
\proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} \citep{R} with the packages
\pkg{glogis}~\Sexpr{packageDescription("glogis")["Version"]},
\pkg{strucchange}~\Sexpr{packageDescription("strucchange")["Version"]} \citep{Zeileis+Leisch+Hornik:2002}, and
\pkg{fxregime}~\Sexpr{packageDescription("fxregime")["Version"]} \citep{Zeileis+Shah+Patnaik:2010},
all of which are available under the General Public License (GPL) from the
Comprehensive \proglang{R} Archive Network (\url{http://CRAN.R-project.org/}). Package \pkg{strucchange} provides
various techniques for inference in structural-change settings which are complemented
with additional maximum-likelihood-based methods by \pkg{fxregime}. Inference for the
generalized logistic model (that can be combined with \pkg{strucchange}/\pkg{fxregime})
is implemented in the new package \pkg{glogis} that also contains the data used, both
the raw monthly price indexes \citep[as provided by][]{oecd} and the seasonally adjusted
monthly inflation rates. The seasonal adjustment was performed in \pkg{X-12-ARIMA}~0.3
(\citealp{X-12-ARIMA}, see also \citealp{Findley+Monsell+Bell:1998}) through the interface
provided by \proglang{gretl}~1.9.5 (\citealp{gretl}, see also \citealp{Smith+Mixon:2006} for a review).

Replication scripts for all analyses are provided within the \pkg{glogis} package: An overview
is provided in \code{demo(package = "glogis")} and the scripts can be launched easily, e.g.,
via \code{demo("Austria", package = "glogis")} etc.

\section*{Acknowledgments}

We are grateful to Jesus Crespo Cuaresma for comments and discussions about
an earlier version of the manuscript, and also to the participants of the workshop
of the Research Platform ``Empirical and Experimental Economics'' for stimulating discussions,
leading to substantial improvements of the paper.

\bibliography{papers}

\end{document}
\SweaveOpts{eval=FALSE}

\clearpage

\begin{appendix}

\section{Data and sales periods}\label{sec:sales}

In addition to Germany, Portugal, Belgium, Italy, Luxembourg, and Spain, there
are a number of other countries where the sales periods are regulated by law.
In France the sale periods time span is regulated by each department. In Greece there
are also two set sales-periods in winter and summer. In Italy the sales periods
are fixed each year by the Chambers of Commerce.  Portugal has two set sales
periods as well, but for two month (January and February as well as Mid-July to
Mid-September). The source for this information is \citet{konseur}. According to
\cite{pltm}, the official dates of the inclusion of these sales periods are:
January 1998 for Germany and Portugal,  January 1999 for Luxembourg and Ireland and 
January 2000 for Belgium,  while Italy and Spain introduced them in January 2001.

\clearpage
\section{Graphs}

Figures~\ref{fig:nochange}--\ref{fig:eastern} depict the data underlying the results
in Section~\ref{sec:results} along with the GL-based means from Table~\ref{tab:results1}.

<<graphs-code>>=
get_data <- function(i) {
  rval <- do.call("merge",
    lapply(i, function(j) results[[j]]$data))
  colnames(rval) <- names(results)[i]
  rval
}

get_fitted <- function(i) {
  rval <- do.call("merge",
    lapply(i, function(j) fitted(results[[j]]$breakpoints, breaks = results[[j]]$nbreaks, type = "mean")))
  colnames(rval) <- names(results)[i]
  rval
}

library("lattice")
get_xyplot <- function(i) {
  nam <- names(results)[i]
  nam[nam == "CzechRepublic"] <- "Czech Republic"
  xyplot(merge(get_data(i), get_fitted(i)),
    screen = rep(nam, 2),
    col = rep(c(1, 4), each = length(i))) 
}
@

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[p!]
\begin{center}
<<graphs-nochange, fig=TRUE, height=6.3, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(2, 4, 8, 14)))
@
\caption{\label{fig:nochange} No change.}

<<graphs-phase1, fig=TRUE, height=4.9, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(1, 6, 7)))
@
\caption{\label{fig:phase1} Phase~1 convergence.}
\end{center}
\end{figure}

\begin{figure}[p!]
\begin{center}
<<graphs-phase2, fig=TRUE, height=6.3, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(9, 12, 13, 16)))
@
\caption{\label{fig:phase2} Phase~2 convergence.}

<<graphs-crisis, fig=TRUE, height=3.5, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(11, 19)))
@
\caption{\label{fig:crisis} Financial crisis.}
\end{center}
\end{figure}

\begin{figure}[p!]
\begin{center}
<<graphs-noneuro, fig=TRUE, height=3.5, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(20, 21)))
@
\caption{\label{fig:noneuro} Non-euro.}

<<graphs-eastern, fig=TRUE, height=9.1, width=7.5>>=
trellis.par.set(theme = standard.theme(color = FALSE))
print(get_xyplot(c(3, 5, 10, 15, 17, 18)))
@
\caption{\label{fig:eastern} Eastern countries.}
\end{center}
\end{figure}

\end{appendix}

\end{document}